\documentclass[11pt]{article}
\usepackage[a4paper,margin=1.05in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{microtype}
\setlength{\emergencystretch}{3em}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}

\title{Predictive Fibres and Critical Oseen Transfer for the Three-Dimensional Navier-Stokes Equations}
\author{Luca Blanchi}
\date{June 28, 2026}

\begin{document}
\maketitle

\begin{abstract}
This note develops an audited presentation-theoretic regularity program for the
three-dimensional incompressible Navier-Stokes equations. The basic object is a
local predictive presentation: incoming and lateral data determine admissible
predictors, while the compatible suitable weak states form an invisible fibre. A
putative singularity is represented by a persistent failure of causal prediction
across parabolic scales.

The current result is a precise conditional reduction, not yet a non-conditional
regularity theorem. The paper proves several structural ingredients and isolates
the remaining analytic burdens as named source theorems. The proved components
include a local critical Oseen transfer estimate in the Serrin range, relaxed
minimax centres, Hilbertian innovation packing, the two-mode null calculation for
the Leray-projected bilinear symbol, exact Beltrami absorption, and exact
two-and-a-half-dimensional reduction. The audit now also gives a detailed
source-contract layer for the quotient packet route and for the pressure/gauge
channel. Certified-new packet rows, representation errors, packet escapes, packet
compatibility, S4 owners, pressure envelopes, pressure representatives, pressure
tail production, pressure shrink/margins, and S8/Oseen owner imports are separated
so that no estimate can borrow from S1, S7, the Oseen production theorem, or the
final global cost statement.

The present version incorporates the latest pressure, force-native,
cutoff-localization, and Oseen owner audit. The pressure-envelope magnitude block
GP1 is closed as a dyadic record dichotomy: after finitely many paid updates the
pressure chart is bounded, while infinitely many strict records give a
non-summable registered pressure escape. The same record mechanism closes the
\(M_p\) coordinate of pressure representative variation and the
pressure-envelope-determined shrink coordinate
\((M_p,\varepsilon_p,\theta_P,s_P)\). A separate finite-state argument closes
invisible drift of the pressure-force pair convention and owner labels. These are
formal source closures, not pressure-tail production estimates.

The Oseen cutoff audit is also sharpened. For a fixed centered dyadic cutoff
family, the complete integrated derivative-weighted cutoff collar is source-closed
by a Hardy domination lemma. In a fixed drift epoch, the dyadic drift-gradient
collar is source-closed by the Serrin exponent, Hardy, and Sobolev interpolation.
These results close real subchannels of the Oseen/S8 interface. They do not close
the full S8 localization theorem, the moving-predictor drift channel, the shared
pressure-force pair, or the native force and carrier source theorems.

Under the remaining source modules, a bad predictive chain cannot persist:
either normal innovation is square-summable by Hilbert packing, the fibre enters a
regular rigid branch, or a carrier/chart/envelope/margin escape is registered.
The article therefore gives a professional conditional proof architecture: all
closed components are proved in detail or reduced to finite ledger identities, while
the remaining blocks are explicitly named, source-scoped, and forbidden from using
the desired conclusion.
\end{abstract}

\section{Purpose}

The incompressible Navier-Stokes equations in three space dimensions are
\[
\partial_t u-\Delta u+(u\cdot\nabla)u+\nabla p=0,
\qquad
\nabla\cdot u=0.
\]
The global regularity problem remains open. Classical local theory studies suitable weak solutions through scale-invariant quantities, pressure decompositions, local energy inequalities, and epsilon-regularity criteria.

This note proposes a complementary organization. Around a putative singular point, one asks whether the core of a parabolic cylinder is causally predictable from incoming and lateral information by a strong local model. The states compatible with the same incoming observables form a fibre. A singular chain is then interpreted as a sequence of scales at which no bounded predictor captures all states in the fibre.

The presentation-theoretic language is used through analytic statements. The paper separates three levels:
\begin{enumerate}[label=(\roman*)]
\item statements proved here;
\item conditional propositions proved from explicitly stated analytic modules;
\item analytic modules used as hypotheses in the modular theorem.
\end{enumerate}
The level of each statement is marked explicitly by its environment: lemma, proposition, theorem, or assumption. The current manuscript is written as an audited reduction. It is meant to preserve the useful structure of the argument while making clear which statements are proved, which are formal bookkeeping closures, and which remain genuine Navier-Stokes source theorems.

\section{Local Setting}

\subsection{Parabolic scaling}

For \(z_0=(x_0,t_0)\), write
\[
Q_R(z_0)=B_R(x_0)\times(t_0-R^2,t_0).
\]
The Navier-Stokes scaling is
\[
u^{(R)}(x,t)=R u(x_0+Rx,t_0+R^2t),
\qquad
p^{(R)}(x,t)=R^2p(x_0+Rx,t_0+R^2t).
\]
A quantity is critical if its normalized value is stable under this scaling.

\subsection{Suitable weak solutions}

A pair \((u,p)\) is a suitable weak solution on \(Q\) if
\[
u\in L^\infty_tL^2_x(Q)\cap L^2_t\dot H^1_x(Q),
\qquad
p\in L^{3/2}_{t,x}(Q),
\]
the equations hold distributionally, and the local energy inequality holds. For every nonnegative \(\phi\in C_c^\infty(Q)\),
\[
\begin{aligned}
&\int |u(x,t)|^2\phi(x,t)\,dx
+2\int_{-\infty}^t\int |\nabla u|^2\phi\,dx\,ds\\
&\le
\int_{-\infty}^t\int
|u|^2(\partial_s\phi+\Delta\phi)
+(|u|^2+2p)u\cdot\nabla\phi\,dx\,ds.
\end{aligned}
\]

\subsection{Pressure cleaning}

Let \(B\subset\mathbb R^3\) be a smooth ball. For \(1<s<\infty\), the local Helmholtz projection \(\mathbb P_B\) is
\[
\mathbb P_B f=f-\nabla \pi_B,
\]
where \(\pi_B\) solves the Neumann problem
\[
\Delta \pi_B=\nabla\cdot f,
\qquad
\partial_n\pi_B=f\cdot n,
\]
with zero average normalization. Standard elliptic estimates give
\[
\|\mathbb P_B f\|_{L^s(B)}+\|\nabla\pi_B\|_{L^s(B)}
\le C_{s,B}\|f\|_{L^s(B)}.
\]

For a local argument, the pressure is decomposed schematically as
\[
p=p_{\mathrm{loc}}+p_{\mathrm{harm}}+p_{\mathrm{tail}}.
\]
The local part is absorbed by \(\mathbb P_B\). The harmonic and nonlocal pieces are not ignored: their contribution is registered by a nonnegative quantity
\[
\operatorname{PressTail}(Q_R,Q_r).
\]
In each estimate below, this term means precisely the residual pressure contribution not controlled by the local projection.

\section{Predictive Presentations}

\subsection{Incoming data}

Fix \(0<\theta<1\) and \(0<\lambda<1\), with \(\theta^2<\lambda\). For \(Q_R(z_0)\), define the incoming region
\[
\mathcal I^-(Q_R)
=
B_R(x_0)\times(t_0-R^2,t_0-\lambda R^2)
\cup
(B_R(x_0)\setminus B_{\theta R}(x_0))\times(t_0-R^2,t_0).
\]
It contains a past slab and a lateral collar, but excludes the final core.

\subsection{Predictors and fibres}

A causal strong predictor on \(Q_R\) is a strong solution \(g\) constructed from data measurable on \(\mathcal I^-(Q_R)\), after the chosen pressure-cleaning procedure. The critical predictor budget is
\[
\|g\|_{L^q_tL^p_x(Q_R)}\le L,
\qquad
\frac2q+\frac3p=1,
\qquad
p>3.
\]
Additional budget components may record pressure-tail size, observable cost, localization cost, and conditioning of the finite observational Gramian.

Let \(\mathcal O\) be a finite family of incoming observables. The invisible fibre associated with \((u,p)\) is the class
\[
\mathsf S^-_{L,\mathcal O}(Q_R)
\]
of suitable weak states \((v,q)\) in \(Q_R\) with budget at most \(L\) and the same incoming observations as \((u,p)\):
\[
\mathcal O(v,q)=\mathcal O(u,p).
\]
Let \(\mathsf P^-_{L,\mathcal O}(Q_R)\) denote the corresponding class of causal strong predictors.

\begin{definition}[Local predictive defect]
Let \(d_{Q_R}\) be a pressure-cleaned distance on the core \(Q_{\theta R}\). The minimax predictive defect is
\[
\mathfrak D_{L,\mathcal O}(Q_R)
=
\inf_{g\in \mathsf P^-_{L,\mathcal O}(Q_R)}
\sup_{(v,q)\in \mathsf S^-_{L,\mathcal O}(Q_R)}
d_{Q_R}((v,q),g).
\]
When the infimum is taken in the closed convex hull of the predictor class, the distance from the relaxed centre to realizable predictors is recorded as a realization defect \(\mathfrak V(Q_R)\).
\end{definition}

A bad predictive chain at \(z_0\) is a nested sequence \(Q_{R_k}(z_0)\), \(R_k\downarrow0\), for which \(\mathfrak D_{L,\mathcal O_k}(Q_{R_k})\) stays bounded below by a fixed positive number.

\section{Critical Local Oseen Transfer}

The following estimate is the most classical analytic ingredient of the program. It is included because it is the mechanism that transports a perturbation around a Serrin-critical strong predictor from one scale to a smaller core.

\begin{lemma}[Critical local Oseen transfer]
Let \(Q_r\Subset Q_R\). Let \(g\) be divergence-free and
\[
g\in L^q_tL^p_x(Q_R),
\qquad
\frac2q+\frac3p=1,
\qquad
p>3.
\]
Let \(r\) be divergence-free and pressure-cleaned, satisfying
\[
\partial_t r-\Delta r
+(g\cdot\nabla)r
+(r\cdot\nabla)g
+\nabla q
=F
\]
in \(Q_R\), distributionally. Choose \(\zeta\in C_c^\infty(Q_R)\) with \(\zeta=1\) on \(Q_r\). Assume that the nonlocal pressure contribution is bounded by \(\operatorname{PressTail}_{R,r}\), and define the incoming-collar cost by
\[
\begin{aligned}
\operatorname{Coll}_{\zeta,g}(r)^2
=&
\sup_{\mathrm{initial\ slab}}\int |r|^2\zeta^2\,dx\\
&+
\int_{Q_R}|r|^2
\left(
|\partial_t\zeta|+|\Delta\zeta|+|\nabla\zeta|^2
\right)\,dx\,dt\\
&+
\int_{Q_R}|g|\,|r|^2|\nabla\zeta|\,dx\,dt.
\end{aligned}
\]
Then
\[
\|r\|_{L^\infty_tL^2_x(Q_r)}
+\|\nabla r\|_{L^2_{t,x}(Q_r)}
+\|r\|_{L^{10/3}_{t,x}(Q_r)}
\]
\[
\le
C
\left(
\operatorname{Coll}_{\zeta,g}(r)
+\|F\|_{L^2_tH^{-1}_x(Q_R)}
+\operatorname{PressTail}_{R,r}
\right),
\]
where \(C\) depends on \(Q_r\Subset Q_R\), on \(p\), and on \(\|g\|_{L^q_tL^p_x(Q_R)}\).
\end{lemma}

\begin{proof}
Test the equation against \(r\zeta^2\). The local pressure projection removes the local pressure contribution; the remaining pressure term is bounded by the definition of \(\operatorname{PressTail}_{R,r}\). We obtain
\[
\frac12\frac{d}{dt}\int |r|^2\zeta^2
+\int |\nabla r|^2\zeta^2
\le
I_{\mathrm{cut}}+I_g+I_F+I_{\mathrm{press}}.
\]
The cutoff term is bounded by the collar part of \(\operatorname{Coll}_{\zeta,g}(r)^2\). The forcing term is estimated by duality:
\[
|\langle F,r\zeta^2\rangle|
\le
\|F\|_{H^{-1}}\|r\zeta^2\|_{H^1}
\le
\frac14\|\zeta\nabla r\|_2^2
+C\|F\|_{H^{-1}}^2
+C\|r\nabla\zeta\|_2^2.
\]
Since \(\nabla\cdot g=0\),
\[
\int (g\cdot\nabla r)\cdot r\zeta^2
=
-\frac12\int |r|^2g\cdot\nabla(\zeta^2),
\]
which is included in the collar cost.

It remains to estimate the stretching term. Using \(\nabla\cdot r=0\),
\[
\int r_i(\partial_i g_j)r_j\zeta^2
=
-\int g_jr_i(\partial_i r_j)\zeta^2
-\int g_jr_ir_j\partial_i(\zeta^2).
\]
The second term is again a collar term. For the first,
\[
\left|
\int g_jr_i(\partial_i r_j)\zeta^2
\right|
\le
\|g\|_{L^p}
\|r\zeta\|_{L^{2p/(p-2)}}
\|\zeta\nabla r\|_{L^2}.
\]
By Gagliardo-Nirenberg,
\[
\|r\zeta\|_{L^{2p/(p-2)}}
\le
C
\|r\zeta\|_{L^2}^{1-3/p}
\|\nabla(r\zeta)\|_{L^2}^{3/p}.
\]
Writing
\[
E(t)=\int |r|^2\zeta^2,
\qquad
D(t)=\int |\nabla r|^2\zeta^2,
\]
and absorbing the cutoff contribution into the collar cost, Young's inequality gives
\[
|I_g^{\mathrm{int}}|
\le
\frac14D(t)+C\|g(t)\|_{L^p}^{q}E(t),
\qquad
q=\frac{2p}{p-3}.
\]
The exponent is exactly the Serrin exponent because \(2/q+3/p=1\).

Thus
\[
\frac{d}{dt}E(t)+\frac12D(t)
\le
C\|g(t)\|_{L^p}^{q}E(t)+B(t),
\]
where \(B(t)\) contains only forcing, pressure-tail, and collar terms. Since \(\|g(t)\|_{L^p}^{q}\in L^1_t\), Gronwall yields
\[
\sup_tE(t)+\int D(t)\,dt
\le
C
\left(
\operatorname{Coll}_{\zeta,g}(r)^2
+\|F\|_{L^2_tH^{-1}_x(Q_R)}^2
+\operatorname{PressTail}_{R,r}^2
\right).
\]
Finally,
\[
\|r\|_{L^{10/3}(Q_r)}
\lesssim
\|r\|_{L^\infty_tL^2_x(Q_r)}^{2/5}
\|\nabla r\|_{L^2(Q_r)}^{3/5}.
\]
This proves the claim.
\end{proof}

\subsection{Dyadic cutoff collars}

The local Oseen lemma is a one-cylinder estimate. Its global use along a candidate
bad chain requires a source theorem for the cutoff collars. This subsection records
the part that is now closed.

Fix an anchored target \(z_*=(x_*,t_*)\), a radius \(R\), and dyadic radii
\[
r_k=2^{-k}R.
\]
Choose once and for all smooth profiles \(\phi\in C_c^\infty(B_b(0))\) and
\(\psi\in C^\infty(\mathbb R)\), with \(0<a<b<1\), such that
\(\phi=1\) on \(B_a(0)\), and \(\psi\) is supported in a fixed past time interval
and equals one on a smaller past time interval. Define
\[
\zeta_k(x,t)=
\phi\!\left(\frac{x-x_*}{r_k}\right)
\psi\!\left(\frac{t_*-t}{r_k^2}\right).
\]
The profile, center, ratios, and time-floor constants are part of the presentation
state. They are not selected after the bad block is observed.

\begin{proposition}[Fixed dyadic cutoff-collar source]
Let
\[
W_k=|\partial_t\zeta_k|+|\Delta\zeta_k|+|\nabla\zeta_k|^2 .
\]
Then, on the fixed outer cylinder,
\[
\sum_{k\ge K} W_k(x,t)
\le
C_W\bigl(|x-x_*|^{-2}+R^{-2}\bigr)
\]
almost everywhere. Consequently, for every residual \(w\) with local energy and
dissipation on the outer cylinder,
\[
\sum_{k\ge K}\int |w|^2 W_k
\le
C
\left(
\int |\nabla w|^2+R^{-2}\int |w|^2
\right).
\]
Thus the integrated derivative-weighted Oseen cutoff collar is a summable source
for this fixed centered dyadic cutoff family.
\end{proposition}

\begin{proof}
The derivative bounds of the fixed profiles give
\[
W_k\le C r_k^{-2}
\bigl(\mathbf 1_{A_k^{\mathrm{lat}}}
      +\mathbf 1_{A_k^{\mathrm{time}}}\bigr),
\]
where the lateral support satisfies
\[
A_k^{\mathrm{lat}}\subset
\{a r_k\le |x-x_*|\le b r_k,\ 0\le t_*-t\le c r_k^2\},
\]
and the time-floor support satisfies
\[
A_k^{\mathrm{time}}\subset
\{|x-x_*|\le b r_k,\ c_0r_k^2\le t_*-t\le c_1r_k^2\}.
\]
For the lateral part, \(r_k\) is comparable to \(|x-x_*|\), and a fixed spatial
point lies in only finitely many dyadic annuli. Hence
\[
\sum_k r_k^{-2}\mathbf 1_{A_k^{\mathrm{lat}}}
\le C |x-x_*|^{-2}.
\]
For the time-floor part, the time condition allows only finitely many dyadic
indices for each fixed \(t<t_*\). On the support, \(|x-x_*|\le br_k\), so
\[
r_k^{-2}\le C |x-x_*|^{-2}
\]
for \(x\ne x_*\). The point \(x=x_*\) is negligible for the pointwise estimate and
is handled by Hardy's inequality in the integrated estimate. A fixed outer cutoff
contributes only a finite \(R^{-2}\) term. This proves the displayed weight
domination.

Let \(\eta_R\) be a fixed spatial cutoff equal to one on the outer ball and
supported in a slightly larger ball. Hardy's inequality applied to \(\eta_R w\)
gives, for a.e. time,
\[
\int |w|^2|x-x_*|^{-2}
\le
C\int |\nabla w|^2 + C R^{-2}\int |w|^2 .
\]
Integrating in time and using the weight domination proves the source estimate.
\end{proof}

\begin{remark}
This proposition closes only the integrated cutoff term
\(\int |w|^2W_k\). It does not close pointwise-in-time collar energy, a
\(\sup_t\) collar ledger, critical annular \(L^3\) localization mass, moving
centers, block-selected cutoff profiles, or the pressure, carrier, force, and
state-motion companions of S8.
\end{remark}

\begin{proposition}[Fixed-drift dyadic drift-collar source]
Assume in addition that, on a fixed merged Oseen epoch, the drift \(g\) and residual
\(w\) are fixed on the outer cylinder and
\[
g\in L^q_tL^p_x,\qquad p>3,\qquad \frac2q+\frac3p=1.
\]
Then
\[
\sum_{k\ge K}\int |g|\,|w|^2|\nabla\zeta_k|<\infty .
\]
More precisely, for every \(\varepsilon>0\),
\[
\sum_{k\ge K}\int |g|\,|w|^2|\nabla\zeta_k|
\le
\varepsilon\int |\nabla w|^2
+C_\varepsilon(1+R^{-2})
\|w\|_{L^\infty_tL^2_x}^2
\|g\|_{L^q_tL^p_x}^{q}.
\]
\end{proposition}

\begin{proof}
The fixed dyadic gradient weights satisfy
\[
\sum_k|\nabla\zeta_k(x,t)|\le C|x-x_*|^{-1}.
\]
Therefore it is enough to control
\[
\int |g|\,|w|^2|x-x_*|^{-1}.
\]
For each time choose \(s\) by
\[
\frac1p+\frac1s+\frac12=1,
\qquad s=\frac{2p}{p-2}.
\]
Hölder's inequality gives
\[
\int |g|\,|w|^2|x-x_*|^{-1}
\le
\|g\|_p\|w\|_s\|w/|x-x_*|\|_2 .
\]
Since \(2<s<6\), Sobolev interpolation gives
\[
\|w\|_s
\le C\|w\|_2^{1-3/p}\|\nabla w\|_2^{3/p}.
\]
The localized Hardy inequality gives
\[
\|w/|x-x_*|\|_2
\le C\|\nabla w\|_2+C R^{-1}\|w\|_2.
\]
Combining these estimates,
\[
\int |g|\,|w|^2|x-x_*|^{-1}
\le
C\|g\|_p\|w\|_2^{1-3/p}
\left(
\|\nabla w\|_2^{1+3/p}
+R^{-1}\|w\|_2\|\nabla w\|_2^{3/p}
\right).
\]
Young's inequality, with \(q=2p/(p-3)\), yields
\[
\int |g|\,|w|^2|x-x_*|^{-1}
\le
\varepsilon\|\nabla w\|_2^2
+C_\varepsilon(1+R^{-2})\|g\|_p^q\|w\|_2^2 .
\]
Integrating in time proves the stated summability.
\end{proof}

\begin{remark}
The fixed-drift hypothesis is essential. If the predictor \(g_k\) and residual
\(w_k=u-g_k\) change at every scale, this proposition cannot be applied until the
drift transitions have been source-closed, finitely merged, or routed to escape or
gate. The general moving-drift collar remains part of the S6/S8/S10/S12 owner
problem.
\end{remark}

\section{Minimax Centres}

The minimax construction used by the predictive framework is a Hilbert-space statement. The PDE content lies in verifying that the chosen predictor classes satisfy the hypotheses.

\begin{lemma}[Relaxed minimax centres]
Let \(\mathcal H\) be a Hilbert space. Let \(P\subset\mathcal H\) be nonempty, bounded, closed, and convex, and let \(S\subset\mathcal H\) be bounded. Define
\[
R(g)=\sup_{v\in S}\|v-g\|_{\mathcal H},
\qquad
g\in P.
\]
Then \(R\) attains its minimum on \(P\).
\end{lemma}

\begin{proof}
The map \(g\mapsto\|v-g\|_{\mathcal H}\) is convex and weakly lower semicontinuous for each \(v\). Hence \(R\), as a supremum of weakly lower semicontinuous functions, is weakly lower semicontinuous. Since Hilbert spaces are reflexive, a bounded closed convex subset is weakly compact. Therefore \(R\) attains its minimum on \(P\).
\end{proof}

\begin{remark}
In the Navier-Stokes application, \(P\) is the weakly closed convex hull of admissible strong predictors at a fixed finite observation budget. The gap between a relaxed minimax centre and an actual strong predictor is exactly the realization defect \(\mathfrak V\). This defect must be paid for in the scale-transfer inequality.
\end{remark}

\section{Innovation Packing}

\subsection{Abstract packing}

Let \(\mathfrak H\) be a Hilbert space of normalized adjoint observables. Suppose that a sequence of normal defects \(\mathcal N_k\in\mathfrak H^*\) has normalized innovative witnesses \(e_k\in\mathfrak H\). The filtration condition is
\[
e_k\perp \mathcal P_{k-1},
\qquad
\mathcal P_k=\overline{\mathcal P_{k-1}+\operatorname{span}\{e_k\}},
\]
so \((e_k)\) is orthonormal on innovative scales.

\begin{lemma}[Hilbertian innovation packing]
Assume that there exist \(\mathcal R\in\mathfrak H^*\), errors \(\rho_k\), and \(c>0\) such that
\[
\langle \mathcal N_k,e_k\rangle
=
\langle \mathcal R,e_k\rangle+\rho_k,
\qquad
\sum_k|\rho_k|^2<\infty,
\]
and
\[
|\langle \mathcal N_k,e_k\rangle|\ge c\|\mathcal N_k\|_{\mathfrak H^*}.
\]
Then
\[
\sum_k\|\mathcal N_k\|_{\mathfrak H^*}^2
\le
C_c\left(\|\mathcal R\|_{\mathfrak H^*}^2+\sum_k|\rho_k|^2\right).
\]
\end{lemma}

\begin{proof}
By Bessel's inequality,
\[
\sum_k|\langle\mathcal R,e_k\rangle|^2
\le
\|\mathcal R\|_{\mathfrak H^*}^2.
\]
Using the coefficient representation,
\[
\sum_k|\langle\mathcal N_k,e_k\rangle|^2
\le
2\sum_k|\langle\mathcal R,e_k\rangle|^2
+2\sum_k|\rho_k|^2.
\]
The witness lower bound then gives the stated estimate.
\end{proof}

\subsection{PDE meaning}

The lemma is abstractly complete. Its PDE use is the global residual representation module: all normalized defects from all scales must be transported into one Hilbert space with square-summable connection, cutoff, gauge, and pressure errors.

\section{Repaired Cost and Quotient Packet Source Contracts}

The audit of the proof is most stringent at the point where Hilbert innovation is
converted into packet estimates. Hilbert orthogonality of the innovation witnesses
does not by itself give a Navier-Stokes packet row estimate. The row family must be
realized in a fixed or finitely merged packet atlas, and every error surrounding the
row estimate must have an owner independent of the row estimate itself.

\subsection{Finite repaired source ledger}

The repaired cost is organized as a finite weighted source expansion
\[
Q_k^{\mathrm{rep}}
=\operatorname{Innov}_k
+\sum_{s\in S_{\mathrm{err}}}w_s\,\operatorname{Err}_{s,k}
+\sum_{x\in S_{\mathrm{esc}}}w_x\,\operatorname{Esc}_{x,k}
+\sum_{u\in S_{\mathrm{upd}}}w_u\,\operatorname{Upd}_{u,k}.
\]
For each source channel exactly one of the following outcomes is legal:
\begin{enumerate}[label=(\roman*)]
\item a summable source theorem;
\item a finite update or finite-atlas merge with registered boundary cost;
\item a named non-summable escape;
\item a gate, meaning entry into a regularity or predictive closure alternative;
\item an explicitly open PDE theorem.
\end{enumerate}
Only the first two outcomes contribute finite error terms on a no-gate tail. A
named non-summable escape exits the no-escape theorem; it is not a finite error in
disguise. An explicitly open PDE theorem records an honest remaining analytic
burden and cannot be used as a closed input to the final proof.

\subsection{The S1 quotient packet row}

For an innovative scale, let
\[
e_k=\frac{d_k-P_{V_{k-1}}d_k}{\|d_k-P_{V_{k-1}}d_k\|},
\qquad
m_k=A_k^*e_k.
\]
In a fixed packet frame write
\[
m_k=\operatorname{Syn}_E(a_k)+\tau_k,
\qquad
a_k=a_k^{\mathrm{old}}+a_k^{\mathrm{cert}}+a_k^{\mathrm{new}}
     +a_k^{\mathrm{tail}}+a_k^{\mathrm{esc}}.
\]
The S1 row theorem is reserved for the certified-new part \(a_k^{\mathrm{new}}\).
Old, certified, recurrent, tail, and escape components are routed to their own
owners before S1 is invoked.

\begin{assumption}[S1 quotient packet row source]
On a fixed or finitely merged packet tail, the certified-new rows satisfy one of the
standard fixed-atlas row criteria: finite overlap, Schur/Gram summability, a
Carleson packet estimate, or a direct row-operator bound. Equivalently, the operator
\[
T_{\mathrm{new}}c=(\langle c,a_k^{\mathrm{new}}\rangle)_k
\]
is bounded on the fixed packet coefficient space. The allowed conclusion is
\[
\sum_k\|a_k^{\mathrm{new}}\|_{\ell^2}^2
\le
C_{\mathrm{S1}}
+C_{\mathrm{S1}}\sum_k\operatorname{ErrPktCompat}_k^{\mathrm{ind}}
+C_{\mathrm{S1}}\sum_k\operatorname{EscPkt}_k,
\]
where the two right-hand side channels have source outcomes independent of this
row theorem.
\end{assumption}

\begin{remark}
This assumption is still one of the central Navier-Stokes PDE tasks. It may not be
inferred from Hilbert Gram-Schmidt orthogonality of the \(e_k\)'s, because the rows
\(a_k^{\mathrm{new}}\) live in scale-dependent packet coordinates with pressure,
cutoff, carrier, gauge, and chart transport errors.
\end{remark}

\subsection{Representation inequality}

Let \(b_k=\operatorname{An}_E(d_k)\) be the fixed-frame coefficient vector of the
registered defect. The packet pairing contract requires
\[
\begin{aligned}
\langle d_k,e_k\rangle_{H_{\mathrm{def}}}
&=\langle b_k,a_k\rangle_{\ell^2}
+\rho_{\mathrm{frame},k}
+\rho_{\mathrm{reg},k}
+\rho_{\mathrm{native},k}.
\end{aligned}
\]
All discrepancies are owned by the repaired ledger. If
\[
\|b_k\|_{\ell^2}\le M_{\mathrm{pkt}},
\]
then Cauchy-Schwarz and Young give
\[
\operatorname{Innov}_k
\le
C_{\mathrm{repr}}\|a_k^{\mathrm{new}}\|_{\ell^2}^2
+C_{\mathrm{repr}}\operatorname{ErrS1repr}_k^{\mathrm{ind}}.
\]
The independent representation error is decomposed as
\[
\begin{aligned}
\operatorname{ErrS1repr}_k^{\mathrm{ind}}
=&\operatorname{ErrPairFrame}_k+\operatorname{ErrPairReg}_k
  +\operatorname{ErrPairNative}_k\\
&+\operatorname{ErrOldPair}_k+\operatorname{ErrCertPair}_k
  +\operatorname{ErrTailPair}_k+\operatorname{ErrPairEnv}_k .
\end{aligned}
\]
The seven owners are, respectively, the packet-frame owner, the registration owner,
the native component owners, the memory/span-gap owner, the certified physical
subtraction owner, the tail owner, and the packet-envelope owner.

\begin{proposition}[S1 representation and escape source contract]
The representation inequality is a closed input to S1 only after every component of
\(\operatorname{ErrS1repr}^{\mathrm{ind}}\) has a legal source outcome independent
of S1. Packet escape is legal on the S1 right-hand side only in the registered
summable case. More precisely,
\[
\begin{aligned}
\operatorname{EscPkt}_k
=&\operatorname{EscFrame}_k+\operatorname{EscCarrier}_k+\operatorname{EscDict}_k
  +\operatorname{EscEnvelope}_k\\
&+\operatorname{EscOwnerSplit}_k+\operatorname{EscRecurrent}_k
  +\operatorname{EscMarginFloor}_k+\operatorname{EscAtlasVar}_k .
\end{aligned}
\]
Each packet escape component must be registered summable, finitely updated or
merged, a named non-summable exit, a gate, or an explicitly open PDE theorem. A
large value of \(\|a_k^{\mathrm{new}}\|_{\ell^2}^2\) alone is not a packet escape.
\end{proposition}

\begin{proof}
This is a bookkeeping theorem in the finite repaired ledger. The representation
terms are precisely the discrepancies between the Hilbert pairing and the fixed
packet pairing, plus old/certified/tail/envelope pairings. By construction none of
these terms contains the certified-new row norm, and each has a primary owner in
S2-S12, S4, S5, the packet-envelope ledger, finite variation, or a named escape.
The escape list records failures of packet data, not large row mass. Therefore the
row estimate and the representation inequality can use these terms without proving
their own right-hand sides.
\end{proof}

\subsection{S4 packet compatibility owner matrix}

The packet compatibility error that may appear beside S1 is the independent part
with the certified-new row mass removed:
\[
\operatorname{ErrPktCompat}_k=\operatorname{ErrPktCompat}_k^{\mathrm{ind}}.
\]
It is decomposed as
\[
\begin{aligned}
\operatorname{ErrPktCompat}_k^{\mathrm{ind}}
=&\operatorname{FrameCompat}_k+\operatorname{RowTailCompat}_k
  +\operatorname{SectorCompat}_k+\operatorname{DictCompat}_k\\
&+\operatorname{OwnerSplitCompat}_k+\operatorname{CertOldCompat}_k\\
&+\operatorname{EnvCompat}_k+\operatorname{MarginPktCompat}_k .
\end{aligned}
\]
The eight primary owner ledgers are:
\[
R_{\mathrm{frame}},\quad R_{\mathrm{tail}},\quad R_{\mathrm{sector}},
\quad R_{\mathrm{dict}},\quad R_{\mathrm{owner}},\quad R_{\mathrm{cert}},
\quad R_{\mathrm{env}},\quad R_{\mathrm{margin}}.
\]
The local S4 estimate on finite legal blocks is
\[
\begin{aligned}
\sum_{k\in B}\operatorname{ErrPktCompat}_k^{\mathrm{ind}}
&\le
C_{\mathrm{S4}}\sum_{k\in B}\bigl(
R_{\mathrm{frame},k}+R_{\mathrm{tail},k}
 +R_{\mathrm{sector},k}+R_{\mathrm{dict},k}\\
&\qquad
 +R_{\mathrm{owner},k}+R_{\mathrm{cert},k}
 +R_{\mathrm{env},k}+R_{\mathrm{margin},k}
\bigr).
\end{aligned}
\]

\begin{proposition}[S4 owner aggregation firewall]
The aggregate symbol \(R_{\mathrm{S4}}\) is only shorthand for the eight owner
ledgers above. It is not an independent source theorem. On an infinite no-gate
no-escape tail, S4 summability follows only if all eight owner families have
summable source outcomes after finitely many certified updates or merges. If one
owner exits by named escape or gate, the S4 summability route exits; if one owner is
only an open PDE theorem, S4 remains conditional.
\end{proposition}

\begin{proof}
The finite-block estimate is the finite sum of the eight local owner estimates.
Passing to an infinite tail is justified only by monotone convergence after each
owner tail has a finite source theorem and all update boundary costs are registered.
There is no ninth analytic estimate hidden in the notation \(R_{\mathrm{S4}}\).
\end{proof}

\subsection{Current packet-accounting status}

The packet-accounting layer is therefore formally closed in the following sense:
\begin{enumerate}[label=(\roman*)]
\item the certified-new row is the only object counted by S1;
\item the representation error has seven independent owners;
\item packet escape has eight named components and cannot be large new-row mass;
\item packet compatibility has eight S4 owners;
\item S4 aggregation is a finite owner sum, not a separate source theorem;
\item no S1 surrounding source may use S1 row Bessel, final \(\mathrm{GQ1}\), or a threshold
chosen after the row is known.
\end{enumerate}
The analytic burden that remains is correspondingly sharp: prove the actual source
outcomes for the representation owners, packet escapes, and eight S4 owner
families; prove the true S1 row Bessel/Carleson theorem for Navier-Stokes packet
rows; and stabilize the packet atlas through the finite-variation and finite-atlas
theorems.

\section{The Bilinear Symbol and Null Fibres}

The next lemma is a finite-dimensional calculation for the Leray-projected quadratic symbol. It explains why large invisible fibres with very small quadratic leakage should be close to special geometric branches, but by itself it treats only pure modes.

\begin{lemma}[Two-mode null classification]
Let
\[
u=a e^{i\xi\cdot x},
\qquad
v=b e^{i\eta\cdot x},
\]
with
\[
a\cdot\xi=0,
\qquad
b\cdot\eta=0.
\]
Set \(k=\xi+\eta\), \(k\ne0\), and consider the symmetrized projected symbol
\[
\mathfrak b((\xi,a),(\eta,b))
=
\mathbb P_k\left[(a\cdot\eta)b+(b\cdot\xi)a\right].
\]
Write
\[
a=\alpha k+A,
\qquad
b=\beta k+B,
\qquad
A\perp k,
\qquad
B\perp k.
\]
Then \(\mathfrak b=0\) if and only if
\[
\alpha B+\beta A=0.
\]
The nonzero solutions fall into the following branches:
\begin{enumerate}[label=(\roman*)]
\item the shear-null branch \(\alpha=\beta=0\);
\item the balanced equal-radius branch \(\alpha\beta\ne0\), which forces \(|\xi|=|\eta|\).
\end{enumerate}
\end{lemma}

\begin{proof}
Since \(a\cdot\xi=0\) and \(b\cdot\eta=0\),
\[
a\cdot\eta=a\cdot(\xi+\eta)=a\cdot k,
\qquad
b\cdot\xi=b\cdot(\xi+\eta)=b\cdot k.
\]
Therefore
\[
(a\cdot\eta)b+(b\cdot\xi)a
=
(a\cdot k)b+(b\cdot k)a
=
|k|^2(\alpha b+\beta a).
\]
Substituting the decompositions gives
\[
\alpha b+\beta a
=
2\alpha\beta k+\alpha B+\beta A.
\]
The projection \(\mathbb P_k\) kills the component parallel to \(k\), so the projected symbol vanishes exactly when \(\alpha B+\beta A=0\).

If \(\alpha=\beta=0\), both polarizations are perpendicular to \(k\), giving the shear-null branch. If exactly one of \(\alpha,\beta\) is zero, then the relation forces one polarization to vanish, so there is no nonzero two-mode interaction. If \(\alpha\beta\ne0\), then \(B=-(\beta/\alpha)A\). The incompressibility condition for \(a\) gives
\[
A\cdot\xi=-\alpha k\cdot\xi.
\]
Since \(A\perp k\), we have \(A\cdot\eta=-A\cdot\xi=\alpha k\cdot\xi\). The incompressibility condition for \(b\) gives
\[
0=b\cdot\eta=\beta k\cdot\eta+B\cdot\eta
=
\beta k\cdot\eta-\frac{\beta}{\alpha}A\cdot\eta
=
\beta(k\cdot\eta-k\cdot\xi).
\]
Thus \(k\cdot\eta=k\cdot\xi\), which is
\[
|\eta|^2+\xi\cdot\eta=|\xi|^2+\xi\cdot\eta.
\]
Hence \(|\xi|=|\eta|\).
\end{proof}

\begin{remark}
The next analytic step is a quantitative packet-level null-graph rigidity theorem: it should promote this exact two-mode calculation to localized parabolic wave packets and identify when a large fibre with small leakage is close to a Beltrami, shear, or carrier-escape branch.
\end{remark}

\section{Rigid Branches}

\begin{proposition}[Beltrami absorption]
Suppose \(U\) is divergence-free and
\[
\nabla\times U=\lambda U.
\]
Then
\[
(U\cdot\nabla)U
=
\nabla\left(\frac{|U|^2}{2}\right).
\]
Thus the quadratic term is pure pressure and is eliminated by Leray projection or by local pressure-cleaning.
\end{proposition}

\begin{proof}
Use the identity
\[
(U\cdot\nabla)U
=
\nabla\left(\frac{|U|^2}{2}\right)
-U\times(\nabla\times U).
\]
If \(\nabla\times U=\lambda U\), then \(U\times(\lambda U)=0\).
\end{proof}

\begin{proposition}[Two-and-a-half-dimensional reduction]
Let
\[
U(x,t)=V(y,t)+W(y,t)n,
\]
where \(y\) belongs to a fixed plane \(\Pi\), \(V(y,t)\in\Pi\), \(n\perp\Pi\), and \(\nabla_\Pi\cdot V=0\). Then the three-dimensional Navier-Stokes system reduces to
\[
\partial_t V+V\cdot\nabla_\Pi V-\Delta_\Pi V+\nabla_\Pi P=0,
\]
\[
\partial_t W+V\cdot\nabla_\Pi W-\Delta_\Pi W=0.
\]
Thus the planar part is a two-dimensional Navier-Stokes solution and the transverse component is a passive transported-diffused scalar.
\end{proposition}

\begin{proof}
All fields are constant in the \(n\)-direction. The in-plane component of the equation is exactly the two-dimensional incompressible Navier-Stokes equation for \(V\). The \(n\)-component contains no pressure gradient and gives the passive scalar equation for \(W\).
\end{proof}

\section{The Remaining Analytic Modules}

We now state the analytic modules still needed by the closed regularity argument.
The previous section fixed the source-contract bookkeeping. The assumptions below
are the remaining scale-invariant PDE assertions.

\begin{assumption}[Predictive epsilon regularity]
For every \(L\) and every admissible pressure-cleaning scheme, there exist \(\varepsilon(L)>0\) and \(0<\theta<1\) such that the following holds. If \((u,p)\) is suitable in \(Q_1\), and if there is a strong predictor \(g\) with
\[
\|g\|_{L^q_tL^p_x(Q_1)}\le L,
\qquad
\frac2q+\frac3p=1,
\qquad
p>3,
\]
such that
\[
\|u-g\|_{X(Q_\theta)}
+\operatorname{PressTail}(Q_1,Q_\theta)
\le\varepsilon(L),
\]
then \(u\) is regular in a smaller cylinder \(Q_{\theta/2}\). Here \(X\) is the pressure-cleaned local energy-critical control norm used in the predictive distance.
\end{assumption}

\begin{assumption}[Critical predictor-map differentiability]
The causal strong solution map \(\Phi\), from incoming data to core predictors, is twice differentiable in the selected critical topology on each bounded predictor class. Quantitatively, for incoming data \(a\) and perturbations \(h\),
\[
\|\Phi(a+h)-\Phi(a)-D\Phi(a)h\|_X
\le
C_L\|h\|_{\mathcal A}^2,
\]
with constants stable under parabolic rescaling and compatible with pressure-cleaning.
\end{assumption}

\begin{assumption}[Global residual representation]
Along every bad predictive chain, there is a normalized Hilbert space \(\mathfrak H\), quasi-isometric identifications of the scale-dependent witness spaces with \(\mathfrak H\), a single residual \(\mathcal R\in\mathfrak H^*\), and square-summable errors \(\rho_k\) such that
\[
\langle\mathcal N_k,e_k\rangle
=
\langle\mathcal R,e_k\rangle+\rho_k
\]
for every innovative normal witness \(e_k\). The errors include pressure tails, cutoffs, holonomy, gauge changes, and finite-window transport errors.
\end{assumption}

\begin{assumption}[Null-graph rigidity for packets]
If a localized parabolic wave-packet family has quadratically small normal leakage on a quantitatively connected interaction graph, then, after removing pressure and gauge directions, it is close in the critical norm to one of the classified branches: Beltrami, two-and-a-half-dimensional shear, or carrier escape. If it is not close to such a branch, it produces a registered normal defect.
\end{assumption}

\begin{assumption}[Carrier localization and collapse]
High-frequency escape either admits a carrier recentering that returns the dominant frequency to order one, or is locally subcritical in every carrier cell, or regenerates critical mass through a registered normal, holonomy, pressure-tail, or localization defect.
\end{assumption}

\begin{assumption}[Quantitative closure alternative]
For every \(L\) and \(\delta>0\), there are \(m=m(L,\delta)\) and \(c=c(L,\delta)>0\) such that the following holds along any \(m\)-step controlled block of a predictive chain. If \(\mathfrak D_k\ge\delta\) throughout the block and all registered normal, holonomy, pressure, carrier, and realization defects in the block are at most \(c\), then either the defect decreases below \(\delta/2\) at the next controlled scale, or the chain enters a rigid branch or a carrier branch covered by the previous assumptions.
\end{assumption}

\begin{assumption}[Completion of the repaired source ledger]
Every source channel in the repaired ledger has one of the legal outcomes listed in
the finite source expansion:
\begin{enumerate}[label=(\roman*)]
\item a summable source theorem;
\item a finite update or finite-atlas merge with registered boundary cost;
\item a named non-summable escape;
\item a regularity gate;
\item an explicitly open PDE theorem.
\end{enumerate}
For the final non-conditional theorem, the fifth alternative must be eliminated.
In particular, the following families require actual Navier-Stokes source
theorems: the eight S4 owner families, the seven \(\operatorname{ErrS1repr}\)
owners, the eight \(\operatorname{EscPkt}\) components, the recurrent/Hilbert-old
alternatives, the S2-S12 transport and component ledgers, the finite-variation
route \(\mathrm{VarEff}\), and the exact E12 floor.
\end{assumption}

\begin{assumption}[Exact floor and finite-atlas closure]
The E12 floor used in the block selection is chosen after all relevant weights,
packet envelopes, margin states, and update costs are fixed. On every no-gate
no-escape tail, either the finite-atlas and finite-variation states stabilize after
finitely many registered updates, or the branch exits by a named chart, packet,
carrier, margin, floor, or component escape. The floor is never recomputed for an
already selected block without paying an explicit transport or update cost.
\end{assumption}

\section{Conditional Newton Transfer}

Under the differentiability module, the Newton bridge becomes a precise conditional proposition.

\begin{proposition}[Conditional saddle Newton bridge]
Assume critical predictor-map differentiability and the local Oseen transfer estimate. Let \(g_k\) be a relaxed minimax predictor at scale \(k\), and write \(u=g_k+w_k\). Then, after passing to the next core scale,
\[
\mathfrak D_{k+1}
\le
C_L\mathfrak D_k^2
+C_L
\left(
\mathfrak K_k^\perp
+\mathfrak H_k^\perp
+\mathfrak F_k^\perp
+\mathfrak V_k
\right)
+\eta_k.
\]
Here \(\mathfrak K_k^\perp\) is the non-absorbed normal quadratic defect, \(\mathfrak H_k^\perp\) the holonomy or gauge defect, \(\mathfrak F_k^\perp\) the pressure/localization defect, \(\mathfrak V_k\) the realization defect of the relaxed centre, and \(\eta_k\) the chosen minimization error.
\end{proposition}

\begin{proof}
The predictable linear part of \(w_k\) is absorbed by the first variation \(D\Phi(a)h\). The tangential quadratic part is absorbed by the second-order correction supplied by the differentiability module. The non-absorbed quadratic part is exactly \(\mathfrak K_k^\perp\); pressure, cutoff, and connection errors give \(\mathfrak F_k^\perp\) and \(\mathfrak H_k^\perp\); and non-realisability of the relaxed centre gives \(\mathfrak V_k\). Comparing the suitable weak solution with the updated strong predictor, the local Oseen estimate bounds the remainder on the smaller core. Since the updated predictor is an admissible competitor for the minimax problem at the next scale, the displayed inequality follows.
\end{proof}

\section{Modular Regularity Theorem}

\begin{theorem}[Audited conditional predictive regularity]
Let \((u,p)\) be a suitable weak solution near \(z_0\). Suppose that the predictive
presentation system at \(z_0\) satisfies the source contracts of the repaired ledger,
the analytic modules stated above, the S1 quotient packet row source, the S7
pressure/gauge source blocks GP1--GP5 with the closed GP1 record and GP2/GP4
coordinate subowners inserted, the post-firewall S6 native-force source contract,
the post-firewall S8 cutoff/localization source contract, and the exact
floor/finite-atlas closure. Assume that the Oseen cutoff collars are taken from the
fixed dyadic family above, so their integrated cutoff part is already source-closed;
assume also that every moving Oseen drift epoch is either finitely merged into a
fixed-drift epoch covered by the fixed-drift collar proposition or has a separate
source-closed transition ledger.
Suppose finally that the minimization errors \(\eta_k\) are summable and that no
source channel needed by the proof is left in the explicitly open PDE category.
Then there is no bad predictive chain at \(z_0\). Consequently, if the predictive
defect is the only obstruction to predictive epsilon regularity, then \(u\) is
regular at \(z_0\).
\end{theorem}

\begin{proof}
Assume that a bad predictive chain exists. Then for some \(\delta_0>0\),
\[
\mathfrak D_k\ge\delta_0
\]
on all sufficiently small scales.

By the global residual representation and the Hilbertian packing lemma, the
innovative normal defects are square-summable. The repaired source ledger and the
S1/S4 firewalls ensure that the surrounding packet, representation, escape,
transport, pressure, cutoff, carrier, margin, and floor errors are either summable
on the no-gate tail, finitely updated before the next epoch, or have already forced
a named escape or gate. The inserted GP1 record theorem and GP2/GP4 coordinate
closures ensure that pressure-envelope growth, \(M_p\)-representative drift,
discrete pair/owner switching, and pressure-envelope shrink degeneration cannot
move invisibly. The remaining GP2--GP5 source blocks must then ensure that residual
representative drift, pressure-tail production, residual shrink/margins,
pressure-force pair mass, native force-use motion, critical localization mass,
carrier tails, moving-drift errors, floor motion, and state variation have single
owners and pre-E12 ledgers before \(243\) and \(265\) are invoked. The fixed dyadic
cutoff lemma removes the integrated Oseen cutoff collar from the hard S8 burden,
and the fixed-drift collar lemma removes the drift-gradient collar on any genuinely
fixed or merged Oseen epoch. Thus on a genuine no-gate no-escape tail, blocks of
arbitrarily small scales have very small average registered defect.

The quantitative closure alternative prevents a persistent defect
\(\mathfrak D_k\ge\delta_0\) on such blocks unless the chain enters a rigid branch
or a carrier branch.

If the chain enters a Beltrami branch, the quadratic term is pressure-cleaned and the Newton transfer gives decay of the predictive defect. If it enters a two-and-a-half-dimensional branch, the reduced two-dimensional system and passive scalar equation provide the strong predictor and again force decay, up to a registered anisotropic defect. If the chain enters carrier escape, the carrier localization module either recenters the scale, makes the contribution locally subcritical, or produces one of the registered defects already controlled by packing.

Thus every legal alternative contradicts the persistence of
\(\mathfrak D_k\ge\delta_0\). Therefore no bad predictive chain exists. The final
implication is exactly predictive epsilon regularity.
\end{proof}

\section{Verified Components and Remaining Work}

The verified pieces are:
\begin{enumerate}[label=(\roman*)]
\item the Serrin-critical Oseen transfer estimate, with collar and pressure-tail terms explicitly paid for;
\item the fixed centered dyadic cutoff-collar source theorem for the integrated
Oseen cutoff weight;
\item the fixed-drift dyadic drift-gradient collar source theorem on fixed or
finitely merged Oseen epochs;
\item the Hilbert-space existence of relaxed minimax centres;
\item the abstract Bessel packing mechanism for innovative normal witnesses;
\item the exact two-mode null classification of the projected bilinear symbol;
\item exact absorption of Beltrami fields and exact reduction of two-and-a-half-dimensional fields.
\end{enumerate}

The formal source-contract layer now additionally closes:
\begin{enumerate}[label=(\roman*)]
\item the quotient packet split into old, certified, new, tail, and escape rows;
\item the representation inequality
\(\operatorname{Innov}_k\le C\|a_k^{\mathrm{new}}\|_{\ell^2}^2
+C\operatorname{ErrS1repr}_k^{\mathrm{ind}}\);
\item the seven-owner decomposition of \(\operatorname{ErrS1repr}\);
\item the eight-component decomposition of \(\operatorname{EscPkt}\), with summable
escape separated from non-summable branch exit;
\item the eight-owner S4 packet-compatibility matrix;
\item the S4 aggregation firewall, which prevents \(R_{\mathrm{S4}}\) from being
used as an independent source theorem;
\item the rule that S1, S4, E12, and final \(\mathrm{GQ1}\) cannot prove the source estimates
that they consume.
\item the S7 pressure/gauge source architecture: only spatial constants are
zero-cost, nonconstant harmonic pressure transfers are pressure-force pair
morphisms, and the \(\operatorname{GaugePress}\) channel is conditionally assembled
from five named source blocks GP1--GP5.
\item the GP1 pressure-envelope magnitude block by dyadic record exhaustion:
eventual bounded pressure chart after finitely many updates, non-summable
\(\operatorname{PressEnvRecEsc}\), or gate/impossible setup;
\item the \(M_p\) coordinate of GP2 pressure representative variation, whose drift
is exactly the GP1 record/update/escape mechanism;
\item the discrete GP2 pair-convention and owner-label coordinates, whose finite
pre-E12 state either stabilizes, pays a non-summable switching ledger, or exits by
fixed-chart illegality;
\item the GP4 pressure-envelope shrink coordinate
\((M_p,\varepsilon_p,\theta_P,s_P)\), so that the route
\(M_p\to\infty\), \(\theta_P\to0\) cannot occur invisibly.
\item the post-firewall S6 native-force source contract, which separates Hilbert
fixed-chart force tails from native force use, native transition, pressure-force
pair, native force-tail, and owner-transition burdens;
\item the post-firewall S8 cutoff/localization normal form, which rejects
\(\operatorname{EscS8}=\operatorname{S8Loc}\) as a source theorem and replaces it
by fixed owner splits, explicit residuals, genuine localization/cutoff escapes,
gates, or impossibility;
\item the Oseen/S8 subsource split for \(\mathrm{Cut}_O\) and fixed-drift
\(\mathrm{DriftCollar}_O\), with all moving-center, moving-drift, pressure,
carrier, and force companions left in their named owners.
\end{enumerate}

The remaining analytic work is concentrated in the following PDE source theorems:
\begin{enumerate}[label=(\roman*)]
\item prove a true S1 row Bessel/Carleson theorem for the certified-new
Navier-Stokes packet rows;
\item prove the packet pairing identity and packet-envelope outcomes for actual
registered Navier-Stokes defects;
\item prove source outcomes for the seven representation-error owners;
\item prove source outcomes or legal exits for the eight packet-escape components;
\item prove source outcomes for all eight S4 owner families: frame, row-tail and
native, carrier-sector, dictionary and component, owner-split, certified-old,
packet-envelope, and margin/floor;
\item prove recurrent and Hilbert-old alternatives through the certified and
recurrent ledgers, without using S1;
\item prove finite-atlas and finite-variation closure through the \(\mathrm{VarEff}\) route;
\item prove the exact E12 floor and its compatibility with packet envelopes, margins,
updates, and block selection;
\item insert the closed GP1 record theorem and the closed GP2/GP4 coordinate
subowners into S7, then prove residual GP2 pressure representative variation, GP3
pressure-tail block production, the remaining GP4 shrink/margin/floor theorem, and
GP5 owner-split import for S8, Oseen, S6, and S11;
\item prove the shared pressure-force-pair source theorem and the S8 pressure/collar
and explicit Oseen-error source outcomes needed by S7;
\item prove the native S6 source theorem: forced-use or zero-certified-use
exhaustion, native \(Y_{\mathrm{force}}\) transition, native force-tail source,
force-owner transition, and the shared pressure-force pair route;
\item prove the full S8 source theorem beyond the integrated cutoff subsource:
critical annular velocity mass, pressure/gauge localization, cutoff normal defect,
carrier imports, predictor tails, recentering transport, containment, and
cutoff/localization state motion;
\item prove S9 carrier source closure and the S10/Oseen moving-owner source theorem,
including pressure pair, force/gauge pair, localization-carrier-cutoff
compatibility, and moving-drift residuals not covered by fixed epochs;
\item prove the S11--E10--E11--E12 shrink-margin-floor source closure and S12/GQ3
finite-state or finite-variation stabilization for all moving owner states;
\item prove critical predictor-map differentiability, global residual
representation, packet-level null-graph rigidity, carrier localization, and the
quantitative closure alternative in the concrete Navier-Stokes topology.
\end{enumerate}

The useful conclusion of the note is therefore a dependency graph with firewalls.
Presentation theory supplies the language of predictors, observables, fibres, costs,
owners, and transfer; Oseen estimates supply the perturbative transport; Hilbert
packing supplies a square-summability mechanism for genuinely new normal defects;
and null geometry isolates the branches where the quadratic term can be invisible.
The next analytic task is to prove that these formal branches exhaust the localized
PDE behaviour at critical scale, with all packet and component source channels
closed before the final global cost statement is invoked.

\appendix

\section{Detailed Audit Ledger: Current Source Contracts}

This appendix records the current fine-grained proof audit. The entries are not a
separate proof of the missing Navier-Stokes PDE source estimates. They are the
source contracts, owner decompositions, and anti-circularity firewalls that the final
proof must respect. The purpose is to keep the article synchronized with the present
state of the proof architecture.

\subsection{L153. S1 quotient packet source contract}

S1 is fixed to the quotient/certified packet route. For every innovative scale,
\[
e_k=\frac{d_k-P_{V_{k-1}}d_k}{\|d_k-P_{V_{k-1}}d_k\|},
\qquad
m_k=A_k^*e_k,
\]
and in the fixed packet frame
\[
m_k=\sum_{\lambda}a_{k,\lambda}\psi_\lambda+\tau_k.
\]
The row is split as
\[
a_k=a_k^{\mathrm{old}}+a_k^{\mathrm{cert}}+a_k^{\mathrm{new}}
     +a_k^{\mathrm{tail}}+a_k^{\mathrm{esc}}.
\]
Only \(a_k^{\mathrm{new}}\) belongs to S1. The S1 source theorem must prove
\[
\sum_k\|a_k^{\mathrm{new}}\|_{\ell^2}^2
\le
C_{\mathrm{S1}}(\mathrm{Envelope},\mathrm{Atlas},\mathrm{Frame})
+C_{\mathrm{S1}}\sum_k\operatorname{ErrPktCompat}_k
+C_{\mathrm{S1}}\sum_k\operatorname{EscPkt}_k,
\]
equivalently boundedness of
\[
T_{\mathrm{new}}c=(\langle c,a_k^{\mathrm{new}}\rangle)_k.
\]
It must also compose with the quotient representation inequality for
\(\operatorname{Innov}_k\). Certified-old, Hilbert-old-uncertified, recurrent, tail,
and escape packet pieces are not counted by S1; they are routed to S4, S5, S6,
S11, S12, \(\mathrm{VarEff}\), escape, or gate.

Remaining burden: prove the actual quotient packet Bessel/Carleson or direct row
operator bound; prove the representation inequality for actual Navier-Stokes
defects; prove packet-frame stability under pressure, Leray, cutoff, gauge, carrier,
force, and chart transitions; prove recurrent physical-lift/certification
alternatives without using S1; prove finite-atlas and finite-variation entry; and
integrate with S4, E12, and final \(\mathrm{GQ1}\).

\subsection{L154. S1 quotient representation inequality}

For a legal S1 scale,
\[
\operatorname{Innov}_k=|\langle d_k,e_k\rangle_{H_{\mathrm{def}}}|^2.
\]
The packet pairing identity must give
\[
\langle d_k,e_k\rangle_{H_{\mathrm{def}}}
=\langle b_k,a_k\rangle_{\ell^2}
+\rho_{\mathrm{frame},k}
+\rho_{\mathrm{reg},k}
+\rho_{\mathrm{native},k},
\]
where \(b_k\) is the registered defect coefficient vector and \(a_k\) is the fixed
packet row of \(A_k^*e_k\). With the owner split of \(a_k\), all old, certified,
tail, and escape pairings are zero in the fixed owner state or charged to owned
repaired errors and escapes. If
\[
\|b_k\|_{\ell^2}\le M_{\mathrm{pkt}},
\]
then Cauchy-Schwarz and Young give
\[
\operatorname{Innov}_k
\le
C_{\mathrm{S1,repr}}\|a_k^{\mathrm{new}}\|_{\ell^2}^2
+C_{\mathrm{S1,repr}}\operatorname{ErrS1repr}_k.
\]
This closes the formal algebra behind the representation inequality, but not the
PDE facts that make the pairing identity and envelope true.

Remaining burden: prove the packet pairing identity for actual registered
Navier-Stokes defects; prove the packet envelope or legal escape/update/gate; prove
the S1 row Bessel/Carleson theorem; prove all representation errors independently
of S1 and final \(\mathrm{GQ1}\).

\subsection{L155. S1 packet pairing and envelope contract}

Inside a fixed packet epoch, the packet analysis map is
\[
\operatorname{An}_E:H_{\mathrm{def}}\to\ell^2(\mathrm{Frame}_E),
\qquad
b_k=\operatorname{An}_E(d_k),
\]
and the synthesis map satisfies
\[
m_k=A_k^*e_k=\operatorname{Syn}_E(a_k)+\tau_k.
\]
The pairing identity is legal only through the fixed analysis/synthesis maps,
canonical registration, and owned frame/native errors:
\[
\langle d_k,e_k\rangle_{H_{\mathrm{def}}}
=\langle \operatorname{An}_E(d_k),a_k\rangle_{\ell^2}
+\rho_{\mathrm{frame},k}
+\rho_{\mathrm{reg},k}
+\rho_{\mathrm{native},k}.
\]
The coefficient envelope is the explicit observable
\[
\operatorname{PktEnv}_k
=\|\operatorname{An}_E(d_k)\|_{\ell^2(\mathrm{Frame}_E)}^2.
\]
S1 may use the representation inequality only when
\(\operatorname{PktEnv}_k\le M_{\mathrm{pkt}}^2\), or when envelope failure is
charged as explicit repaired envelope error, finite update, named packet/native/
carrier escape, chart escape, or gate.

Forbidden moves: choosing \(b_k\) after the fact; deriving the envelope from finite
energy at critical normalization; changing registration, frame, dual frame, or
observation map after \(e_k\); hiding pressure, force, cutoff, carrier, localization,
memory, gauge, or normal-defect mismatch in a packet Bessel term.

Remaining burden: construct actual Navier-Stokes packet charts supplying
\(\operatorname{An}_E\), \(\operatorname{Syn}_E\), and canonical registration; prove
envelope outcomes; prove frame residual and native mismatch source outcomes; prove
finite-atlas stabilization and S1 row Bessel/Carleson.

\subsection{L156. S1 row Bessel source criterion}

On a fixed or merged S1 tail atlas, the certified-new row family
\[
\mathcal A_{\mathrm{new}}=\{a_k^{\mathrm{new}}:\operatorname{Innov}_k>0\}
\]
must satisfy at least one genuine row source criterion:
\begin{enumerate}[label=(\roman*)]
\item finite-overlap source;
\item Schur or Gram source;
\item Carleson packet source;
\item direct row-operator source.
\end{enumerate}
Equivalently, \(T_{\mathrm{new}}\) is bounded on the fixed packet coefficient space.
The proof may not infer this from Hilbert Gram-Schmidt orthogonality of \(e_k\) in
\(H_{\mathrm{def}}\). If no allowed row source criterion holds, the legal outcomes
are packet/frame/carrier escape, finite-atlas or finite-variation failure, certified
update or recurrent lift, packet-envelope escape/update, gate, or an explicitly open
S1 PDE theorem.

Remaining burden: prove one allowed criterion for actual Navier-Stokes
certified-new rows; prove fixed packet frame, carrier sector, dictionary, and
envelope stabilization; remove recurrent/Hilbert-old rows through S5/S12 without
using S1; prove independent source outcomes for surrounding errors.

\subsection{L157. S1 surrounding-error independence firewall}

An error term near the S1 row theorem is admissible only if it has a primary owner in
S2-S12, \(\mathrm{VarEff}\), GQ3, or a named escape; its definition excludes
\(\|a_k^{\mathrm{new}}\|^2\), the unknown \(T_{\mathrm{new}}\) norm, and final
\(\mathrm{GQ1}\); and its source theorem does not use the S1 row estimate.

\(\operatorname{ErrPktCompat}_k\) is legal only as packet compatibility with
\(\operatorname{NewRowSize}_k=\|a_k^{\mathrm{new}}\|^2\) removed. It may contain
frame, row-tail, sector, dictionary, owner-split, certified-old, envelope, and
margin errors owned by non-S1 channels. It may not contain certified-new row mass,
recurrent packet mass before the recurrent alternatives are assigned, or a packet
chart change chosen after \(e_k\).

\(\operatorname{ErrS1repr}_k\) is split into frame pairing, registration pairing,
native pairing, old-pair, certified-pair, and tail-pair owners. \(\operatorname{EscPkt}_k\)
is legal only for named packet-chart failures: unavailable frame, carrier-window
exit, dictionary transport failure, envelope exit, owner-split failure, recurrent
state failure, margin/floor degeneration, or finite-atlas/variation failure.

Consequently the S1 row theorem plus the representation inequality is non-circular
only after these surrounding terms have source outcomes independent of S1.

\subsection{L158. S4 packet compatibility source contract}

S4 is the source channel for the S1-independent packet compatibility error
\[
\operatorname{ErrPktCompat}_k^{\mathrm{ind}}.
\]
On a fixed or merged packet epoch,
\[
\begin{aligned}
\operatorname{ErrPktCompat}_k^{\mathrm{ind}}
=&\operatorname{FrameCompat}_k+\operatorname{RowTailCompat}_k
  +\operatorname{SectorCompat}_k+\operatorname{DictCompat}_k\\
&+\operatorname{OwnerSplitCompat}_k+\operatorname{CertOldCompat}_k\\
&+\operatorname{EnvCompat}_k+\operatorname{MarginPktCompat}_k.
\end{aligned}
\]
The owner matrix routes frame and row-tail to S4/S2/S3, sector to S9/carrier,
dictionary to S6-S9/S3/S2, owner split and certified-old to S5/S11/S12, envelope to
the packet-envelope owner, and margin/floor to S11/E10-E12/S12. The source target is
\[
\sum_{k\in B}\operatorname{ErrPktCompat}_k^{\mathrm{ind}}
\le C_{\mathrm{S4}}R_{\mathrm{S4}}(B),
\]
with all repaired-cost weights fixed before E12.

Excluded from S4: \(\operatorname{NewRowSize}\), the unknown \(T_{\mathrm{new}}\)
norm, recurrent-new mass before the recurrent alternatives, post-innovation chart
choices, S1 row Bessel, final \(\mathrm{GQ1}\), and final absence of bad blocks.

\subsection{L159. S4 frame compatibility owner}

In a fixed packet chart with frame \(\mathrm{Frame}_E\), maps
\(\operatorname{An}_E,\operatorname{Syn}_E\), dual frame, and bounds \(A_E,B_E\),
standard frame stability gives
\[
\operatorname{FrameCompat}_k^{\mathrm{fix}}
\le C_{\mathrm{fr}}(\operatorname{FrameErr}_k+\operatorname{RowTail}_k).
\]
For accepted frame transitions, `BuildTrans=OK' and router S2/S4 must first route
packet-frame and registration pieces, giving
\[
\operatorname{FrameCompat}_k^{\mathrm{trans}}
\le C_{\mathrm{fr}}^{\mathrm{trans}}
(\operatorname{PktFrameTrans}_k+\operatorname{RegFrameTrans}_k+\operatorname{RowTail}_k).
\]
Any term proportional to \(\delta_k\|a_k^{\mathrm{new}}\|^2\), or depending on
\(T_{\mathrm{new}}\), is not an S4 frame error. It must be absorbed by a fixed or
merged frame before S1, handled by finite variation or finite update, routed to
escape/gate, or left open.

The block target is
\[
\sum_{k\in B}\operatorname{FrameCompat}_k
\le C_{\mathrm{S4,frame}}R_{\mathrm{frame}}(B).
\]

\subsection{L160. S4 row-tail/native owner}

In
\[
m_k=A_k^*e_k=\operatorname{Syn}_E(a_k)+\tau_k,
\]
the residual \(\tau_k\) has exactly four legal outcomes: exact zero, predeclared
coefficient tail, registered native/transport tail, or escape/gate. If a discarded
component is admissible in the fixed packet chart and is not old, certified, tail,
or escape by the owner map, it belongs to \(a_k^{\mathrm{new}}\), not to \(\tau_k\).

The owner size is
\[
\operatorname{RowTailOwner}_k
=\operatorname{RowTailCoeff}_k+\operatorname{RowTailNative}_k
+\operatorname{RowTailTransport}_k+\operatorname{RowTailReg}_k.
\]
The source targets are
\[
\sum_{k\in B}\operatorname{RowTailCompat}_k
\le C_{\mathrm{S4,tail}}\sum_{k\in B}\operatorname{RowTailOwner}_k,
\]
and, when the packet envelope is legal,
\[
\sum_{k\in B}\operatorname{RowTailPair}_k
\le C_{\mathrm{pair,tail}}\sum_{k\in B}\operatorname{RowTailOwner}_k.
\]
Native pressure, force, cutoff/localization, carrier, chart, normal, memory, and
certificate tails retain their native owners.

\subsection{L161. S4 sector/carrier owner}

Carrier depth, enlargement width, threshold, recentering list, and sector labels are
chart data fixed before \(e_k\). Sector mismatch has exactly one legal outcome:
inside fixed sector, predeclared carrier tail, accepted carrier transport, or
carrier/packet/chart escape or gate. In-sector certified-new mass cannot be hidden
in sector compatibility.

The sector owner is
\[
\operatorname{SectorOwner}_k
=\operatorname{CarPkt}_k+\operatorname{CarTail}_k+\operatorname{CarTrans}_k
+\operatorname{CarPress}_k+\operatorname{CarLoc}_k+\operatorname{CarTheta}_k,
\]
and
\[
\sum_{k\in B}\operatorname{SectorCompat}_k
\le C_{\mathrm{S4,sector}}\sum_{k\in B}\operatorname{SectorOwner}_k.
\]
This is an S9/carrier theorem, not a proof of S1 row Bessel or carrier summability.

\subsection{L162. S4 dictionary/component owner}

Packet dictionary labels are not free coordinates. They are a finite product of
component charts owned by S2, S3, S6, S7, S8, S9, S11, and S12. The dictionary owner
is
\[
\begin{aligned}
\operatorname{DictOwner}_k
=&\operatorname{ForceDict}_k+\operatorname{GaugeDict}_k
  +\operatorname{CutLocDict}_k+\operatorname{CarrierDict}_k\\
&+\operatorname{RegNativeDict}_k+\operatorname{MarginDict}_k.
\end{aligned}
\]
Fixed dictionaries have zero dictionary loss. Accepted component-owned mismatch
satisfies
\[
\sum_{k\in B}\operatorname{DictCompat}_k
\le C_{\mathrm{S4,dict}}\sum_{k\in B}\operatorname{DictOwner}_k.
\]
Certified-new in-dictionary mass remains \(a_k^{\mathrm{new}}\). Nonconstant
pressure/gauge transfers, cutoff changes, carrier changes, and registration/native
changes keep their primary component owners. Dictionary drift cannot replace the
finite-atlas/variation route.

\subsection{L163. S4 owner-split state theorem}

The split
\[
a_k=a_k^{\mathrm{old}}+a_k^{\mathrm{cert}}+a_k^{\mathrm{new}}
     +a_k^{\mathrm{tail}}+a_k^{\mathrm{esc}}
\]
is a deterministic verdict of the fixed E6-E12 effective state, packet state,
recurrent alternatives, and S1 firewall. Every coefficient has exactly one verdict.
In a fixed owner state, \(\operatorname{OwnerSplitCompat}_k^{\mathrm{fix}}=0\).
For accepted owner perturbations,
\[
\sum_{k\in B}\operatorname{OwnerSplitCompat}_k
\le C_{\mathrm{S4,owner}}\sum_{k\in B}\operatorname{OwnerSplitLedger}_k,
\]
where
\[
\begin{aligned}
\operatorname{OwnerSplitLedger}_k
=&\operatorname{MemOwner}_k+\operatorname{AtomIdOwner}_k
  +\operatorname{RankCoeffOwner}_k+\operatorname{ForceOwner}_k\\
&+\operatorname{MarginOwner}_k+\operatorname{FloorOwner}_k
  +\operatorname{RecurrentOwner}_k+\operatorname{TailOwner}_k.
\end{aligned}
\]
Certified-new mass remains in \(a_k^{\mathrm{new}}\). Hilbert-old but uncertified
components go to S5, S12, update, escape, or gate. Recurrent components must pass
through the recurrent alternatives before S1.

\subsection{L164. S4 certified-old physical subtraction}

A certified-old verdict is not physical subtraction. To remove a component from S1,
the proof must exhibit a certified predictor
\[
G_A(c_k)=G_{\mathrm{base}}+\sum_i c_{k,i}G_i
\]
with identity-preserving atom match, canonical coefficients, rank and margin control,
component conventions for force, gauge, cutoff, carrier, and native norms, and
nonlinear defect recomputation
\[
\begin{aligned}
\operatorname{Def}_{NS}(G_A(c_k))
=&\partial_tg_A-\Delta g_A\\
&+\operatorname{div}(g_A\otimes g_A)+\nabla\pi_A.
\end{aligned}
\]
The certified-old ledger is
\[
\begin{aligned}
\operatorname{CertOldLedger}_k
=&\operatorname{CertAtomId}_k+\operatorname{CertCoeffRank}_k
  +\operatorname{CertDefUse}_k+\operatorname{CertForceUse}_k\\
&+\operatorname{CertGaugePair}_k+\operatorname{CertCutLoc}_k
  +\operatorname{CertCarrier}_k\\
&+\operatorname{CertNativeTail}_k+\operatorname{CertMarginFloor}_k.
\end{aligned}
\]
Legal certified-old epochs satisfy
\[
\sum_{k\in B}\operatorname{CertOldCompat}_k
\le C_{\mathrm{S4,cert}}\sum_{k\in B}\operatorname{CertOldLedger}_k.
\]
Hilbert-old is not certified-old subtraction; forced-use cost is paid per use; and
nonconstant pressure/gauge transfers use their registered pressure-force owner.

\subsection{L165. S4 packet-envelope owner}

The packet coefficient envelope
\[
\operatorname{PktEnv}_k
=\|\operatorname{An}_E(d_k)\|_{\ell^2(\mathrm{Frame}_E)}^2
\]
has exactly four legal outcomes:
\begin{enumerate}[label=(\roman*)]
\item bounded chart \(\operatorname{PktEnv}_k\le M_{\mathrm{pkt}}^2\);
\item explicit repaired error \(\operatorname{ErrPktEnv}_k=\operatorname{PktEnv}_k\)
with pre-E12 weight;
\item finite envelope update;
\item envelope escape or gate.
\end{enumerate}
The envelope owner is
\[
\operatorname{EnvOwner}_k
=\operatorname{EnvFrame}_k+\operatorname{EnvNative}_k+\operatorname{EnvCarrier}_k
+\operatorname{EnvMargin}_k+\operatorname{EnvUpdate}_k,
\]
and legal envelope excess satisfies
\[
\sum_{k\in B}\operatorname{EnvCompat}_k
\le C_{\mathrm{S4,env}}\sum_{k\in B}\operatorname{EnvOwner}_k.
\]
\(M_{\mathrm{pkt}}\) is chart data fixed before \(e_k\), S1, and E12. Envelope
failure cannot hide certified-new row mass, and \(\operatorname{ErrS1repr}\) may
use the envelope only after this owner theorem supplies a legal outcome.

\subsection{L166. S4 margin and floor owner}

The margin owner is
\[
\begin{aligned}
\operatorname{MarginOwner}_k
={}&\operatorname{ShrinkOwner}_k+\operatorname{GraphMarginOwner}_k\\
&+\operatorname{CertMarginOwner}_k+\operatorname{PacketEnvMarginOwner}_k\\
&+\operatorname{FloorOwner}_k+\operatorname{BoundaryOwner}_k.
\end{aligned}
\]
\(\operatorname{MarginPktCompat}\) has exactly four legal outcomes: fixed positive
margin, registered margin comparison, finite update/floor recompute, or escape/gate.
In a fixed positive chart, \(\operatorname{MarginPktCompat}_k^{\mathrm{fix}}=0\).
For accepted margin comparisons,
\[
\sum_{k\in B}\operatorname{MarginPktCompat}_k
\le C_{\mathrm{S4,margin}}\sum_{k\in B}\operatorname{MarginOwner}_k.
\]
Shrink stacks, graph/certificate thresholds, packet-envelope thresholds, floor
sources, block lengths, and boundary data are fixed before S1/S4 use. If any change,
the old epoch ends unless an explicit transport theorem preserves it with registered
loss.

\subsection{L167. S4 owner aggregation firewall}

The aggregate
\[
R_{\mathrm{S4}}(B)=\sum_{k\in B}
(R_{\mathrm{frame},k}+R_{\mathrm{tail},k}+R_{\mathrm{sector},k}
+R_{\mathrm{dict},k}+R_{\mathrm{owner},k}+R_{\mathrm{cert},k}
+R_{\mathrm{env},k}+R_{\mathrm{margin},k})
\]
is only shorthand for eight owner ledgers. It is not an independent source theorem.
On finite legal blocks, the eight local inequalities add to
\[
\sum_{k\in B}\operatorname{ErrPktCompat}_k^{\mathrm{ind}}
\le C_{\mathrm{S4}}R_{\mathrm{S4}}(B).
\]
On an infinite no-gate no-escape tail, S4 summability follows only if all eight owner
families have summable source outcomes after finitely many certified updates/merges.
If any owner exits by escape or gate, the route exits. If any owner remains an open
PDE theorem, S4 remains conditional and cannot be used as a closed input to S1 or
E12.

\subsection{L168. S1 representation and packet escape source contract}

The independent representation error is
\[
\begin{aligned}
\operatorname{ErrS1repr}_k^{\mathrm{ind}}
=&\operatorname{ErrPairFrame}_k+\operatorname{ErrPairReg}_k
  +\operatorname{ErrPairNative}_k\\
&+\operatorname{ErrOldPair}_k+\operatorname{ErrCertPair}_k
  +\operatorname{ErrTailPair}_k+\operatorname{ErrPairEnv}_k.
\end{aligned}
\]
The packet escape split is
\[
\begin{aligned}
\operatorname{EscPkt}_k
=&\operatorname{EscFrame}_k+\operatorname{EscCarrier}_k+\operatorname{EscDict}_k
  +\operatorname{EscEnvelope}_k\\
&+\operatorname{EscOwnerSplit}_k+\operatorname{EscRecurrent}_k
  +\operatorname{EscMarginFloor}_k+\operatorname{EscAtlasVar}_k.
\end{aligned}
\]
Representation owners must be supplied independently of S1 by S2-S12, S4, S5, and
packet-envelope owners. Packet escape is legal on the S1 right-hand side only when it
is registered as a summable repaired escape with pre-E12 weight. A non-summable
named escape exits the no-escape branch; it is not a finite S1 error.

The combined S1 surrounding source is
\[
R_{\mathrm{sur}}=R_{\mathrm{S4}}+R_{\mathrm{repr}}+R_{\mathrm{esc}},
\]
where \(R_{\mathrm{S4}}\) is legal only through L167, \(R_{\mathrm{repr}}\) through
this contract, and \(R_{\mathrm{esc}}\) only for registered summable packet escapes.

\subsection{L169. S1 \texorpdfstring{\(\operatorname{ErrPairFrame}\)}{ErrPairFrame} owner}

The first representation-error owner is
\[
\operatorname{ErrPairFrame}_k=|\rho_{\mathrm{frame},k}|^2.
\]
It is the fixed-frame pairing discrepancy after registration and native
discrepancies have been removed. It is generated only by fixed analysis/synthesis
duality defects, dual-frame comparison, row residual \(\tau_k\), accepted
packet-frame transport, or accepted registration/frame pullback comparison.

On a bounded-envelope frame epoch,
\[
\sum_{k\in B}\operatorname{ErrPairFrame}_k
\le C_{\mathrm{S1,pair,frame}}R_{\mathrm{pair,frame}}(B),
\]
Here \(R_{\mathrm{pair,frame}}\) is the \(M_{\mathrm{pkt}}^2\)-weighted sum of the
frame-error, row-tail, packet-frame transport, registration-frame transport, and
finite frame-update costs. The estimate is legal only after the packet-envelope
owner has supplied bounded chart, repaired envelope, finite update, escape, gate, or
open-PDE outcome.

This owner forbids proving frame pairing by S1 row Bessel, final \(\mathrm{GQ1}\),
the unknown \(T_{\mathrm{new}}\) norm, or by selecting the frame after \(e_k\) is
known. Missing envelope or moving frame exits through update, envelope owner, frame
escape, gate, or open theorem.

\subsection{S7 pressure/gauge source architecture}

The S7 pressure/gauge channel is now decomposed into five subowners:
\[
\operatorname{GaugeObs}_k
=\operatorname{GaugeGrad}_k+\operatorname{GaugePress}_k
 +\operatorname{GaugeCut}_k+\operatorname{GaugeShrink}_k
 +\operatorname{GaugePair}_k .
\]
The exact zero-cost fibre is
\[
\operatorname{Gauge}_0=\{c(t)\}.
\]
Every nonconstant harmonic pressure transfer is a pressure-force pair morphism, not
a gauge identity. In the notation of the audit, the pair defect is
\[
\rho_k^{\mathrm{press-force}}
=\|\nabla\psi_k\|_{Y_{\mathrm{force}}}
 +\operatorname{PressTailSize}(\psi_k)
 +\operatorname{GaugeLoss}_k+\operatorname{CutoffPressureLoss}_k.
\]
The local modulus is closed only in the direction
\[
\operatorname{Target}_{\mathrm{gauge}}(B)
\le C_{\mathrm{gauge}}\operatorname{Obs}_{\mathrm{gauge}}(B).
\]
This is not a source theorem. The opposite direction needed for block selection is
provided only by component ledgers satisfying the \(243\) production hypothesis.

\subsection{GaugePress conditional source assembly}

The pressure-tail/envelope/oscillation subchannel \(\operatorname{GaugePress}\) is
conditionally assembled from five blocks:
\begin{description}[leftmargin=3.2cm,style=nextline]
\item[GP1.]
Pressure-envelope source closure. The harmonic envelope magnitude is now closed by
the dyadic record theorem below. It enters an eventual bounded pressure chart after
finitely many record updates, pays the non-summable pressure-envelope record
escape, opens a gate, or is impossible because the pressure representative was not
legal before block selection. This does not produce the pressure-tail observable in
GP3.
\item[GP2.]
Pressure representative variation. The canonical representative is fixed only inside
a fixed physical chart. Across chart changes, pressure representatives must have a
source ledger, a source-closed owner split, finite update, finite variation/Cauchy/
finite atlas, non-summable escape, gate, or impossibility.
\item[GP3.]
Pressure-tail block production. On every legal pressure block,
\[
\operatorname{Obs}_{\mathrm{GaugePress}}(B)
\le \Phi_{\mathrm{press}}(R_{\mathrm{press}}(B)),
\qquad
Q_B^{\mathrm{rep}}\ge w_{\mathrm{press}}R_{\mathrm{press}}(B),
\]
or the pressure tail is assigned by a stored finite split to source-closed owners.
If a block supremum is used, GP3 includes the required sup-to-ledger estimate.
\item[GP4.]
Pressure shrink and margin import. Bounded harmonic envelope may be used through
inner shrink only after \(\theta_P\), \(s_P\), \(\mu_P\), \(\mu_{\mathrm{gauge}}\),
the E10 shrink stack, the E11 margin state, the block length, and the E12 floor data
are fixed, source-controlled, finitely updated, escaped, gated, or impossible.
\item[GP5.]
Cross-channel owner import. Any pressure piece crossing S7, S8, the Oseen channel,
S6, or S11 must have a stored disjoint owner split. In particular, the
\(\mathrm{PressPair}_O\) term needs both pressure/gauge and collar/localization
owners unless it is paid as an explicit Oseen error. The
\(\mathrm{GaugeForceCmp}_O\) term may use a short force/gauge route only after
pressure and cutoff subpieces are assigned.
\end{description}

These five hypotheses imply the following conditional source theorem. If GP1--GP5
hold with weights, tolerances, shrink data, square/unsquare conventions, and owner
splits fixed before E12, then \(\operatorname{GaugePress}\) has exactly one of seven
legal outcomes:
\begin{enumerate}[label=(\roman*)]
\item zero spatial-gauge pressure;
\item a summable primary \(\operatorname{GaugePress}\) ledger;
\item a stored split to source-closed owners;
\item finite update or finite-atlas merge;
\item named non-summable escape;
\item gate;
\item impossibility by fixed-chart/source legality.
\end{enumerate}
Thus \(\operatorname{GaugePress}\) is conditionally ready for S7 and \(265\), but it
is not yet a nonconditional source theorem. The only zero-cost gauge identity is
\(c(t)\); the closures below are source alternatives, not free gauge identities.

\subsection{Closed GP1 pressure-envelope record theorem}

The pressure-envelope part of GP1 is no longer merely a named source obligation. It
has a complete finite-state proof. For every legal scale \(k\), let
\[
P_k=\operatorname{EnvPress}_{R_k}(h_k)<\infty
\]
be the harmonic pressure envelope of the fixed pressure representative. Define the
dyadic level
\[
\nu_k=\min\{n\in\mathbb N:P_k\le 2^n\},
\]
with \(\nu_k=0\) when \(P_k\le1\). The pressure chart stores a record level
\(N_k\), hence a threshold \(M_p(k)=2^{N_k}\), before block selection. A strict
record event is
\[
\nu_k>N_{k-1}.
\]
At such a scale the old chart has failed and the ledger records
\[
\operatorname{PressEnvRecEsc}_k=1.
\]
Only after that boundary is paid may future scales use the updated record
\(N_k=\nu_k\). The record-breaking scale is not retroactively inside the old
bounded pressure chart.

\begin{proposition}[GP1 pressure-envelope source closure]
On every no-gate candidate tail with legal pressure representatives, exactly one of
the following alternatives holds.
\begin{enumerate}[label=(\roman*)]
\item There are finitely many strict dyadic pressure records. Then there exist
\(K_*\) and \(M_p=2^{N_*}<\infty\) such that, for all \(k\ge K_*\),
\[
P_k\le M_p,
\]
and every later finite legal block \(B\) satisfies
\[
\operatorname{PEnv}(B):=\sup_{j\in B}P_j\le M_p.
\]
The last record enlargement is a finite update, after which all dependent pressure
shrink, margin, block-length, and floor data must be recomputed before the next
block is judged.
\item There are infinitely many strict dyadic pressure records. Then
\[
\sum_k\operatorname{PressEnvRecEsc}_k=\infty.
\]
The branch exits by a named non-summable pressure-envelope escape. This is not a
small pressure error and not a pressure-tail production theorem.
\item At some scale the pressure representative, harmonic tail, spatial-average
normalization, or envelope threshold is not legal before block selection. Then the
branch exits by pressure-chart failure, pressure gate, CKN gate, or impossibility.
\end{enumerate}
No fourth GP1 envelope-magnitude outcome is legal.
\end{proposition}

\begin{proof}
The sequence \((\nu_k)\) takes values in the nonnegative integers. If the set of
strict record times is infinite, then the record rule assigns one unit of
\(\operatorname{PressEnvRecEsc}\) at every such time, so its total sum is infinite.
If the set of strict record times is finite, the running maximum stabilizes at a
finite integer \(N_*\). Hence \(\nu_k\le N_*\) on the tail, so \(P_k\le2^{N_*}\) for
every late scale. Taking a supremum over a finite legal block gives the block
envelope bound. If the pressure data are not fixed before the block is known, the
record sequence is not a legal chart coordinate, and the transition router has only
escape, gate, or impossibility outcomes. These cases exhaust the possibilities for
an integer record sequence.
\end{proof}

This theorem replaces the weaker excess observable
\([\,\operatorname{PEnv}(B)/M_p-1\,]_+\), which could in principle be positive on
infinitely many blocks but summable. Strict dyadic records cannot drift upward
invisibly: finite records give a fixed chart, and infinite records give
non-summable escape. The theorem uses no final \(\mathrm{GQ1}\), no pressure-tail
smallness, and no after-the-fact threshold choice.

\subsection{Closed GP2 envelope-coordinate owner}

The pressure representative state contains, among other coordinates,
\[
\operatorname{PressRepState}_k
=(\mathrm{geom},\mathrm{cutoff},p_{\mathrm{loc}},h^0,M_p,\Theta_P,
\mathrm{PairConv},\mathrm{Owner}_P).
\]
After GP1, the \(M_p\) coordinate is no longer an open representative-variation
problem. Split
\[
\Delta\operatorname{PressRep}_k
=\Delta\operatorname{EnvRec}_k+\Delta\operatorname{PressRepRes}_k,
\]
where \(\Delta\operatorname{EnvRec}_k=0\) inside the stored record chart and
\[
\Delta\operatorname{EnvRec}_k=\operatorname{PressEnvRecEsc}_k
\]
at a strict record boundary.

\begin{proposition}[GP2 envelope-coordinate closure]
The \(M_p\) coordinate of \(\operatorname{PressRepState}\) has exactly one legal
source outcome: eventual stabilization after finitely many pressure record updates,
non-summable \(\operatorname{PressEnvRecEsc}\), or pressure-chart gate/impossibility.
In the stable case, the \(M_p\) coordinate contributes no further pressure
representative variation.
\end{proposition}

\begin{proof}
This is the GP1 record dichotomy applied to the threshold coordinate. If records are
finite, \(M_p=2^{N_*}\) is eventually fixed. If records are infinite, every
threshold change is the already registered non-summable pressure-envelope escape.
If the coordinate is chosen after seeing the block, the fixed-chart rules are
violated and the transition is not an accepted representative comparison.
\end{proof}

The residual GP2 theorem is therefore reduced to
\[
\begin{aligned}
\Delta\operatorname{PressRepRes}
={}&\Delta\mathrm{Geom}+\Delta\mathrm{Cutoff}
+\Delta\mathrm{HarmonicNonconstant}\\
&+\Delta\mathrm{ShrinkMargin}
+\Delta\mathrm{PairConvention}+\Delta\mathrm{Owner}
+\Delta\mathrm{OseenS8}.
\end{aligned}
\]
The envelope-threshold drift is not allowed to reappear inside this residual.

\subsection{Closed GP2 pair-convention and owner-label owner}

The pair convention and owner template are finite pre-E12 presentation data. An
accepted discrete pressure-force owner template is a finite object
\[
\begin{aligned}
D_k=(&\mathrm{PairMode},\mathrm{SquareMode},\mathrm{Owner}_{\nabla},
\mathrm{Owner}_{\mathrm{press}},\mathrm{Owner}_{\mathrm{cut}},\\
&\mathrm{Owner}_{\mathrm{shrink}},\mathrm{Owner}_{\mathrm{rem}},
\mathrm{WeightTemplate},\mathrm{ToleranceTemplate}).
\end{aligned}
\]
Define the switch ledger
\[
\operatorname{PairOwnerSwitch}_k=\mathbf 1_{D_k\ne D_{k-1}}.
\]

\begin{proposition}[GP2 discrete pair/owner stabilization]
For a finite registered pair/owner alphabet, exactly one of the following holds.
\begin{enumerate}[label=(\roman*)]
\item \(D_k\) changes finitely many times and is eventually fixed. Then the discrete
pair convention and owner-label coordinates contribute no further GP2 variation.
\item \(D_k\) changes infinitely many times and
\[
\sum_k\operatorname{PairOwnerSwitch}_k=\infty.
\]
The branch exits by a non-summable owner-update ledger, chart escape, or gate.
\item \(D_k\) is selected after the pressure transfer or after the bad block is
known. This violates fixed presentation legality and is impossible, except as an
escape or gate.
\end{enumerate}
\end{proposition}

\begin{proof}
After the budget alphabet is fixed, \(D_k\) takes values in a finite set. A sequence
in a finite set either changes finitely many times or infinitely many times. Finite
changes give an eventual constant state. Infinite changes make the indicator
\(\operatorname{PairOwnerSwitch}_k\) equal to one on infinitely many indices, so the
switch ledger is non-summable. An after-the-fact owner choice is not a sequence of
stored chart states and is rejected by the transition router.
\end{proof}

This closes only the discrete drift. Once the template has stabilized, the analytic
pressure-force pair mass still has to be proved by the selected owner: S7, S6, S8,
S10/Oseen, S11, explicit repaired error, escape, or gate.

\subsection{Closed GP4 pressure-envelope shrink coordinate}

Fix the pressure allowance \(\varepsilon_p>0\) before block selection and before
E12. For a bounded pressure chart \(M_p\), the harmonic pressure shrink coordinate is
\[
\theta_P(M_p)=\min\left\{\frac14,
\left(\frac{\varepsilon_p}{C_PM_p}\right)^{2/5}\right\},
\qquad
s_P=\left\lceil\log_2\frac1{\theta_P}\right\rceil.
\]
The pressure-envelope shrink state is
\[
\Theta_P^{\mathrm{env}}=(M_p,\varepsilon_p,\theta_P,s_P).
\]

\begin{proposition}[GP4 pressure-envelope shrink-coordinate closure]
For every no-gate tail with legal pressure representatives and fixed
\(\varepsilon_p\), the coordinate \(\Theta_P^{\mathrm{env}}\) has exactly one legal
outcome.
\begin{enumerate}[label=(\roman*)]
\item If GP1 has finitely many pressure records, then \(M_p<\infty\) is eventually
fixed, \(\theta_P>0\), \(s_P<\infty\), and this pressure-envelope shrink coordinate
contributes no further \(\operatorname{GaugeShrink}\) transition cost.
\item If GP1 has infinitely many pressure records, then the recomputation indicator
for \(\Theta_P^{\mathrm{env}}\) is non-summable on the same subsequence as
\(\operatorname{PressEnvRecEsc}\).
\item If \(M_p\), \(\varepsilon_p\), or the pressure representative is not fixed
before block selection, the branch exits by chart escape, pressure gate, CKN gate,
or impossibility.
\end{enumerate}
\end{proposition}

\begin{proof}
The only variable in the displayed formula for \(\theta_P\) is \(M_p\), since
\(\varepsilon_p\) and \(C_P\) are fixed theorem data. Finite GP1 records make
\(M_p\) eventually finite and fixed, hence the formula gives a fixed positive
\(\theta_P\) and finite \(s_P\). Infinite GP1 records force infinitely many
pressure-envelope shrink recomputations and therefore the same non-summable record
escape. If the data are not fixed before selection, the shrink is an after-the-fact
choice and is illegal.
\end{proof}

Consequently the degeneration
\[
M_p\to\infty,\qquad \theta_P\to0,\qquad s_P\to\infty
\]
cannot occur at invisible cost. What remains open in GP4 is the rest of the shrink
stack, margin positivity, block containment, E12 floor compatibility, and source
closure for S11/S12 owners.

\subsection{Registered-escape firewall}

The repaired ledger uses named escape streams, but an escape stream is not allowed
to be merely a channel renamed as its own failure. For a nonnegative channel
\(Z_k\), the dichotomy
\[
\sum_k Z_k<\infty
\qquad\hbox{or}\qquad
\sum_k Z_k=\infty
\]
does not by itself prove a source theorem. Otherwise every source channel would
close tautologically by setting \(\operatorname{EscZ}_k=Z_k\).

A registered escape used in the final proof must satisfy the following firewall.
\begin{enumerate}[label=(\roman*)]
\item It has independent failure meaning: chart, owner, construction,
containment, realization, floor, packet, pressure/gauge, force, localization,
carrier, Oseen, or gate data fail.
\item The label, observable, threshold, owner, and repaired weight are fixed before
E12 and before block selection.
\item The escape route is disjoint from the summable-error route.
\item The proof supplies a positive lower bound, monotone/Carleson/packing
contradiction, actual divergence from the failure mechanism, gate, or impossibility
theorem.
\item The proof does not use final \(\mathrm{GQ1}\), future \(243\) block
smallness, or the consumer theorem.
\item If the failure cannot be made into genuine escape, it is recorded as an
impossible branch rather than as a source theorem.
\end{enumerate}

This firewall is now applied to S6, S8, S10/Oseen, S7, S9, S11, S12, and explicit
residual routes. In particular, \(\operatorname{EscS8}=\operatorname{S8Loc}\),
\(\operatorname{ForceUseEsc}=\operatorname{ForceUse}\), and
\(\operatorname{ErrO}=\operatorname{Obs}_O\) are not source theorems without an
independent failure mechanism.

\subsection{S6 native-force source contract}

The force channel is split into a Hilbert fixed-chart part and a native force part:
\[
\mathrm{S6}
=\mathrm{S6}_{\mathrm{HilbFixed}}+\mathrm{S6}_{\mathrm{Native}},
\]
where
\[
\mathrm{S6}_{\mathrm{HilbFixed}}
=\mathrm{Innov}_{\mathrm{force,new}}
 +\mathrm{ForceHilbErr}
 +\mathrm{ForceTail}_H,
\]
and
\[
\mathrm{S6}_{\mathrm{Native}}
=\mathrm{ForceUse}
 +\mathrm{ForceYTrans}
 +\mathrm{ForcePressPair}
 +\mathrm{ForceTail}_{\mathrm{native}}
 +\mathrm{ForceOwner}_{\mathrm{native}}.
\]

The Hilbert fixed-chart part may use the Bessel/innovation machinery only inside
the fixed Hilbert chart. It does not control native \(Y_{\mathrm{force}}\) motion.
The native part is controlled only by the following source alternatives:
\begin{enumerate}[label=(\roman*)]
\item fixed or merged Hilbert epochs plus summable native force cost;
\item a force-use source theorem or a zero-certified-use exhaustion theorem;
\item finite update or finite atlas merge;
\item mergeable finite-variation, Cauchy, monotone, or finite-atlas drift;
\item named non-summable escape satisfying the registered-escape firewall;
\item gate;
\item impossible branch.
\end{enumerate}

The zero-certified-use route has a narrow meaning. If the selected exact/no-force
branch uses no nonzero certified forced atom, then forced-use mass is moved out of
\(\mathrm{ForceUse}\). The burden then shifts to exact atom identity, span-gap,
packet-tail, C1, native-tail, owner-transition, and escape/gate ledgers. It does
not eliminate S7/S8/S10/Oseen/S11 pressure-force or cutoff burdens.

The current native subowners are:
\begin{enumerate}[label=(\roman*)]
\item \(\mathrm{ForceUse}\): either original forced-use source closure or
zero-certified-use exhaustion;
\item \(\mathrm{ForceYTrans}\): native \(Y_{\mathrm{force}}\) comparison and
variation, not merely Hilbert \(H_{\mathrm{force}}\) comparison;
\item \(\mathrm{ForcePressPair}\): the shared pressure-force pair theorem with
pressure, force, gauge, cutoff, shrink, S8, Oseen, S11, and S12 pieces assigned
once;
\item \(\mathrm{ForceTail}_{\mathrm{native}}\): native finite-window, certificate,
localization, carrier, old-force, chart-tail, and owner-state tail routes;
\item \(\mathrm{ForceOwner}_{\mathrm{native}}\): owner-label, atom-identity,
coefficient-rank, native-use-state, pair-state, force-tail-state, and owner-boundary
routes.
\end{enumerate}

Thus S6 is formally organized but not yet a nonconditional source theorem. It can be
used in the modular theorem only after the selected native alternatives are proved
or routed to finite update, genuine escape, gate, or impossibility.

\subsection{S8 cutoff/localization post-firewall normal form}

After the S8 primary owners assigned elsewhere are removed, the active localization
channel is
\[
\begin{aligned}
\mathrm{S8Loc}_{\mathrm{active},k}
=&\mathrm{LocVel}_{\mathrm{active},k}
 +\mathrm{LocPress}_{\mathrm{active},k}
 +\mathrm{LocGauge}_{\mathrm{active},k}
 +\mathrm{LocNorm}_{\mathrm{active},k}\\
&+\mathrm{LocCar}_{\mathrm{active},k}
 +\mathrm{LocPred}_{\mathrm{active},k}
 +\mathrm{LocCtr}_{\mathrm{active},k}
 +\mathrm{LocContain}_{\mathrm{active},k}\\
&+\mathrm{CutLocState}_{\mathrm{active},k}
 +\mathrm{ErrS8Residual}_k .
\end{aligned}
\]
The term \(\mathrm{CutLocState}_{\mathrm{active}}\) records cutoff profile,
localization profile, annulus, center, transport, theta, and floor state motion not
already zero in the fixed anchored chart or owned by S11/S12. The residual term is
zero only when every local piece has a named owner; otherwise it is an explicit
pre-E12 repaired-cost stream.

On a fixed anchored chart, the local localization modulus is closed:
\[
\mathrm{Target}_{\mathrm{loc}}(B)
\le
C_{\mathrm{loc}}\mathrm{Obs}_{\mathrm{loc}}(B).
\]
This local estimate is not source summability of \(\mathrm{Obs}_{\mathrm{loc}}\).
The completed S8 theorem must prove exactly one of:
\begin{enumerate}[label=(\roman*)]
\item direct summability of \(\mathrm{S8Loc}_{\mathrm{active}}\);
\item import from already source-closed owners through a fixed disjoint split;
\item finite update, finite merge, or finite state motion;
\item mergeable finite-variation, Cauchy, monotone, or finite-atlas localization
drift;
\item genuine non-summable \(\mathrm{LocEsc}\) or \(\mathrm{CutEsc}\) satisfying
the registered-escape firewall;
\item gate;
\item impossible branch.
\end{enumerate}
No eighth outcome is legal.

The integrated derivative-weighted part of \(\mathrm{LocVel}\) associated with
Oseen \(\mathrm{Cut}_O\) is now closed for the fixed dyadic cutoff family. The
remaining \(\mathrm{LocVel}\) burden includes critical annular velocity/CKN mass,
pointwise-in-time collar ledgers, \(\sup_t\) ledgers, moving centers, and
commutators measured in stronger native norms. The other S8 subowners remain as
follows:
\begin{enumerate}[label=(\roman*)]
\item \(\mathrm{LocPress}\) and \(\mathrm{LocGauge}\) require S7 or shared
pressure-force pair source closure with no double counting;
\item \(\mathrm{LocNorm}\) requires normal-defect/cutoff-commutator source closure
or an explicit residual theorem;
\item \(\mathrm{LocCar}\) requires S9 carrier source closure and a fixed S8/S9
priority split;
\item \(\mathrm{LocPred}\) requires S3/native predictor-tail routing or explicit
localization residual closure;
\item \(\mathrm{LocCtr}\) is zero only in the anchored chart, otherwise it needs a
transport-back theorem or exits by update, escape, chart escape, or gate;
\item \(\mathrm{LocContain}\) is zero only under fixed positive containment,
shrink, and cylinder-nesting data, otherwise it routes to S11/S12, escape, or gate;
\item \(\mathrm{CutLocState}\) requires S12/GQ3 finite-state or finite-variation
closure, or a genuine cutoff/localization escape.
\end{enumerate}

\subsection{S7/S8/Oseen owner firewall}

Oseen production is now registered on legal blocks:
\[
\operatorname{Obs}_O(B)\le\Phi_O(R_O(B)),
\qquad
Q_B^{\mathrm{rep}}\ge w_OR_O(B),
\]
where
\[
\begin{aligned}
R_O={}&\mathrm{InitO}+\mathrm{Cut}_O+\mathrm{DriftCollar}_O
       +\mathrm{PressPair}_O\\
     &+\mathrm{Force}_O+\mathrm{GaugeForceCmp}_O+\mathrm{LocCarCut}_O .
\end{aligned}
\]
This production theorem is not source summability of the Oseen owners. The pressure
pieces inside \(R_O\) may be imported only after the receiving S7, S8, S6, S11, or
explicit Oseen-error source theorem is closed.

Two Oseen subchannels now have genuine source routes under stated hypotheses. First,
if \(\mathrm{Cut}_O\) uses the fixed centered dyadic family above, its integrated
derivative-weighted cutoff part is source-closed by Hardy and local
energy/dissipation. Second, if the Oseen drift and residual are fixed on a merged
epoch, the drift-gradient collar is source-closed by the fixed-drift Serrin-Hardy
estimate. These closures are narrow: they do not apply to moving centers,
block-selected cutoff profiles, pointwise collar energies, critical annular CKN
mass, or moving predictor drifts.

The remaining Oseen pieces are still owner obligations. \(\mathrm{PressPair}_O\)
requires both pressure/gauge and collar/localization owners unless it is paid as an
explicit Oseen pressure-pair error. \(\mathrm{Force}_O\) requires native
\(L^2_tH^{-1}_x\) force ownership; the cutoff companion created by testing against
\(w\zeta^2\) is not force. \(\mathrm{GaugeForceCmp}_O\) requires a stored
pressure-force pair defect with pressure, force, gauge, cutoff, and shrink pieces
assigned once. \(\mathrm{LocCarCut}_O\) must be split into localization, carrier,
pressure/gauge, force, packet, recentering, containment, cutoff, and explicit-error
pieces before block selection.

Likewise, S8 cutoff/localization readiness is not S8 source closure. The pressure,
gauge, collar, carrier, recentering, containment, and state pieces used by
\(\operatorname{GaugePress}\), \(\operatorname{PressPair}_O\),
\(\operatorname{GaugeForceCmp}_O\), and \(\operatorname{LocCarCut}_O\) must be
summable, imported from source-closed owners, finitely updated, genuinely escaped,
gated, or impossible in their own owners. No pressure tail, cutoff pressure loss,
shrink/margin loss, force-use mass, carrier tail, localization cost, or
pressure-force pair mass may be counted twice.

\subsection{Current unresolved proof obligations}

The detailed ledger above leaves the following genuinely analytic tasks:
\begin{enumerate}[label=(\roman*)]
\item prove source outcomes for the frame, row-tail/native, transport, carrier,
dictionary, owner-split, certified-old, packet-envelope, margin, and floor ledgers
appearing in S4;
\item prove the remaining \(\operatorname{ErrS1repr}\) owners: registration, native,
old-pair, certified-pair, tail-pair, and envelope-pair;
\item prove all eight \(\operatorname{EscPkt}\) components have registered summable,
finite-update, named-exit, gate, or eliminated open-PDE outcomes;
\item prove actual packet pairing identity, packet-envelope outcomes, and a true S1
row Bessel/Carleson or direct row-operator bound for Navier-Stokes rows;
\item prove recurrent and Hilbert-old alternatives without using S1;
\item prove the native S6 theorem: force-use or zero-certified-use exhaustion,
native \(Y_{\mathrm{force}}\) transition, native force-tail source closure,
force-owner transition, Hilbert fixed-chart globalization, and the shared
pressure-force pair route;
\item prove residual S7 pressure/gauge source closure: GP2 representative variation
beyond the closed envelope coordinate, GP3 pressure-tail production, the remaining
GP4 shrink/margin/floor theorem, gauge-gradient, gauge-cutoff, gauge-shrink, and
gauge-pair;
\item prove the shared pressure-force-pair source theorem, including gradient
representative movement, pressure tail, gauge representative loss, gauge shrink
margin loss, cutoff pressure loss, fixed owner split, and genuine non-tautological
escape semantics;
\item prove the full S8 theorem beyond the fixed dyadic cutoff collar: critical
annular \(\mathrm{LocVel}\), \(\mathrm{LocPress}\), \(\mathrm{LocGauge}\),
\(\mathrm{LocNorm}\), \(\mathrm{LocCar}\), \(\mathrm{LocPred}\),
\(\mathrm{LocCtr}\), \(\mathrm{LocContain}\), \(\mathrm{CutLocState}\), and
\(\mathrm{ErrS8Residual}\);
\item prove S9 carrier source closure and all S8/S9 annular-carrier priority splits;
\item prove S10/Oseen source closure beyond the closed fixed subchannels:
moving-drift epochs, pressure pair, force/pair, localization-carrier-cutoff
compatibility, explicit Oseen errors, and no double counting with S6/S7/S8/S9/S11;
\item prove S11--E10--E11--E12 shrink-margin-floor source closure, exact floor
positivity, and compatibility with envelopes, margins, updates, and block
selection;
\item prove S12/GQ3 finite-atlas, finite-variation, Cauchy, monotone,
state-update, escape, or gate alternatives for all moving owner states;
\item close predictor differentiability, global residual representation,
packet-level null-graph rigidity, carrier localization, and the quantitative closure
alternative in the concrete Navier-Stokes topology;
\item assemble final \(\mathrm{GQ1}\) only after all consumed source theorems have
already been proved or legally routed to gate, escape, finite update, or
impossibility.
\end{enumerate}

\begin{thebibliography}{99}

\bibitem[ABP]{ABP}
D. Albritton, T. Barker, and C. Prange,
\emph{Epsilon regularity for the Navier-Stokes equations via weak-strong uniqueness},
arXiv:2211.16188.

\bibitem[CKN]{CKN}
L. Caffarelli, R. Kohn, and L. Nirenberg,
\emph{Partial regularity of suitable weak solutions of the Navier-Stokes equations},
Communications on Pure and Applied Mathematics 35 (1982), 771-831.

\bibitem[ESS]{ESS}
L. Escauriaza, G. Seregin, and V. Sverak,
\emph{\(L_{3,\infty}\)-solutions of Navier-Stokes equations and backward uniqueness},
Russian Mathematical Surveys 58 (2003), 211-250.

\bibitem[Leray]{Leray}
J. Leray,
\emph{Sur le mouvement d'un liquide visqueux emplissant l'espace},
Acta Mathematica 63 (1934), 193-248.

\bibitem[Serrin]{Serrin}
J. Serrin,
\emph{On the interior regularity of weak solutions of the Navier-Stokes equations},
Archive for Rational Mechanics and Analysis 9 (1962), 187-195.

\bibitem[Tao]{Tao}
T. Tao,
\emph{Quantitative bounds for critically bounded solutions to the Navier-Stokes equations},
preprint, 2019.

\bibitem[Wolf]{Wolf}
J. Wolf,
\emph{On the local pressure of the Navier-Stokes equations and related systems},
Advances in Differential Equations 17 (2012), 305-338.

\end{thebibliography}

\end{document}
